---
title: "The Poetry of Resistance: A Cross-Linguistic Text Mining Analysis"
subtitle: "Mahmoud Darwish (Arabic) vs. Amiri Baraka (English)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: flatly
    code_folding: hide
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 6, fig.align = "center",
  dev = if (requireNamespace("ragg", quietly = TRUE)) "ragg_png" else "png",
  dpi = 150
)
```

<div class="project-header">
<div class="team-card">
<h4>Team</h4>
<ul>
<li><span class="member-name">Abdullah Tadmuri</span></li>
<li><span class="member-name">Maiheliya Maimaitimin</span></li>
<li><span class="member-name">Tone Varberg Sabri</span></li>
</ul>
<div class="course-line">Text Mining Final Project &middot; UC3M &middot; 2025&ndash;2026</div>
</div>
<div class="stats-grid">
<div class="stat-card">
<div class="stat-number">28</div>
<div class="stat-label">Poems Analyzed</div>
</div>
<div class="stat-card">
<div class="stat-number">10</div>
<div class="stat-label">Techniques Used</div>
</div>
<div class="stat-card">
<div class="stat-number">2</div>
<div class="stat-label">Languages</div>
</div>
<div class="stat-card">
<div class="stat-number">14</div>
<div class="stat-label">Poems per Poet</div>
</div>
</div>
</div>

# 1. Introduction

Comparing poets who write in different languages requires text mining techniques
that can bridge linguistic boundaries. This study takes on that challenge by
analyzing two canonical resistance poets from distinct traditions: **Mahmoud
Darwish** (1941--2008), the Palestinian national poet who wrote in Arabic, and
**Amiri Baraka** (1934--2014), the African-American poet and activist who wrote
in English.

Both poets use verse as a tool for collective resistance, though they address
different struggles. Darwish's poetry centers on Palestinian dispossession, exile,
and the longing for homeland, while Baraka's work confronts anti-Black racism,
celebrates Black identity, and calls for revolutionary change.

The central question of this analysis is: **How do these two resistance poets
compare in their emotional expression, lexical diversity, and structural patterns?**
A bilingual corpus poses methodological challenges, since Arabic and English words
cannot be directly compared. We address this by selecting techniques that either
produce language-independent numeric outputs (NRC emotion scores, TTR, sparsity) or
run separately per poet for visual side-by-side comparison (word frequency, bigram
networks, TF-IDF, sentiment arcs, pairwise correlation).

Our corpus contains **14 poems per poet** (28 total), drawn from different periods
of each poet's career. Darwish's poems are kept in their original Arabic, with
full diacritics normalization applied. We apply ten text mining techniques:
NRC sentiment analysis, radar charts, poem-emotion heatmaps, sentiment arcs,
TF-IDF, lexical diversity (TTR, MATTR), sparsity, word frequency, bigram
networks, and pairwise word correlation.


# 2. Setup and Data

## 2.1 Packages

```{r packages}
library(tidyverse)
library(tidytext)
library(textdata)
library(fmsb)
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(wordcloud)
library(reshape2)
library(igraph)
library(ggraph)
library(widyr)
library(RColorBrewer)
library(scales)
library(knitr)
library(ggwordcloud)
library(patchwork)
```

## 2.2 Load Poem Data

```{r load-data}
source("darwish_dataset.R")
source("Baraka_dataset.R")

# Harmonize: add poem_id to Baraka, add English poet name to both
baraka_poems <- baraka_poems %>%
  mutate(poem_id = row_number(),
         poet_en = "Amiri Baraka")

darwish_poems <- darwish_poems %>%
  mutate(poet_en = "Mahmoud Darwish")

# Combined tibble for cross-poet numeric comparisons
all_poems <- bind_rows(
  darwish_poems %>% select(poet_en, poem_id, poem, text),
  baraka_poems  %>% select(poet_en, poem_id, poem, text)
)

cat("Darwish poems:", nrow(darwish_poems), "\n")
cat("Baraka poems:", nrow(baraka_poems), "\n")
```

## 2.3 Stopword Preparation

Arabic stopwords must be normalized with the same function used on the poems
(diacritics removal, alif/ya/ta-marbuta normalization). We also add common Arabic
function words missing from the default list and filter out onomatopoeia fragments
from Baraka's poems.

```{r stopwords}
# Arabic normalization (same as clean_poem in darwish_dataset.R)
normalize_ar <- function(text) {
  text <- stringr::str_remove_all(text, "[\\u064B-\\u065F\\u0670]")
  text <- stringr::str_replace_all(text, "ـ", "")
  text <- stringr::str_replace_all(text, "[إأآٱ]", "ا")
  text <- stringr::str_replace_all(text, "ى", "ي")
  text <- stringr::str_replace_all(text, "ة", "ه")
  text <- stringr::str_replace_all(text, "ؤ", "و")
  text <- stringr::str_replace_all(text, "ئ", "ي")
  text
}

# Normalize the default Arabic stopwords
ar_stops_base <- normalize_ar(stopwords::stopwords("ar", source = "misc"))

# Extended custom Arabic stopwords (common function words missing from default)
ar_stops_custom <- normalize_ar(c(
  "على", "أنا", "يا", "أن", "ما", "لا", "إلى", "هذا", "هذه", "ذلك",
  "التي", "الذي", "كان", "كانت", "عن", "بين", "حين", "بعد", "قبل",
  "فوق", "تحت", "حتى", "إذا", "لكن", "ولكن", "أو", "ثم", "كي",
  "لكي", "منه", "منها", "فيه", "فيها", "عليه", "عليها", "إنه", "إنها",
  "أنت", "أنتِ", "نحن", "هم", "لي", "لنا", "لهم", "بها", "به",
  "كل", "بعض", "كيف", "أين", "متى", "لماذا", "هل", "قد", "سوف",
  "كأن", "مثل", "أيها", "وراء", "أمام", "إلا", "أما",
  "ولم", "لكنني", "ولكنني", "كنت", "كنا", "كان", "ليس",
  "الذين", "اللذين", "واو", "فاء"
))

ar_stops <- unique(c(ar_stops_base, ar_stops_custom))
cat("Total Arabic stopwords (normalized):", length(ar_stops), "\n")

# Custom Baraka stopwords (onomatopoeia fragments)
baraka_custom_stops <- c("bu", "de", "daaaa", "bap", "boo", "bup",
                         "pu", "daaaa", "rrrrrrrrrrrrrrr", "rrrrrrrr",
                         "tuhtuhtuhtuhtuhtuhtuhtuhtuhtuh", "aggh",
                         "whoooo", "whooooooooooooooooooooo")

# Unified English stopwords (tidytext + stopwords package + custom noise)
en_stops <- unique(c(stop_words$word, stopwords::stopwords("en"), baraka_custom_stops))
```

## 2.4 Load NRC Multilingual Emotion Lexicon

The NRC Emotion Lexicon (Mohammad & Turney, 2013) maps words in 100+ languages to
8 emotions (anger, anticipation, disgust, fear, joy, sadness, surprise, trust) plus
positive/negative sentiment. Because both Arabic and English words are mapped to the
same emotion categories, this serves as the bridge for cross-language comparison.

```{r nrc-lexicon}
nrc_raw <- read_tsv(
  "NRC-Emotion-Lexicon/NRC-Emotion-Lexicon-ForVariousLanguages.txt",
  show_col_types = FALSE
)

emotion_cols <- c("anger", "anticipation", "disgust", "fear",
                  "joy", "negative", "positive", "sadness",
                  "surprise", "trust")

# English NRC lexicon
nrc_english <- nrc_raw %>%
  select(`English Word`, all_of(emotion_cols)) %>%
  pivot_longer(all_of(emotion_cols),
               names_to = "sentiment", values_to = "value") %>%
  filter(value == 1) %>%
  select(word = `English Word`, sentiment) %>%
  distinct()

# Arabic NRC lexicon
nrc_arabic <- nrc_raw %>%
  select(Arabic, all_of(emotion_cols)) %>%
  pivot_longer(all_of(emotion_cols),
               names_to = "sentiment", values_to = "value") %>%
  filter(value == 1) %>%
  select(word = Arabic, sentiment) %>%
  distinct()

cat("English NRC entries:", nrow(nrc_english), "\n")
cat("Arabic NRC entries:", nrow(nrc_arabic), "\n")
```

## 2.5 Corpus Overview

```{r corpus-overview}
corpus_stats <- all_poems %>%
  mutate(
    word_count = str_count(text, "\\S+"),
    line_count = str_count(text, "\n") + 1
  ) %>%
  group_by(poet_en) %>%
  summarise(
    poems = n(),
    total_words = sum(word_count),
    mean_words = round(mean(word_count)),
    median_words = round(median(word_count)),
    total_lines = sum(line_count),
    .groups = "drop"
  )

kable(corpus_stats,
      col.names = c("Poet", "Poems", "Total Words", "Mean Words",
                     "Median Words", "Total Lines"),
      caption = "Corpus summary statistics")
```


# 3. Analysis

## 3.1 NRC Sentiment and Emotion Analysis

This technique assigns emotion labels to every word in each poem using the NRC
Emotion Lexicon. Because the lexicon provides parallel entries for Arabic and
English words mapped to the same emotions, we can directly compare emotional
profiles across languages. We tokenize each poet's poems separately, join to the
appropriate language lexicon, count emotion matches, and normalize by total matched
words per poem to account for length differences.

```{r sentiment-tokenize}
# Tokenize and join to NRC -- Darwish (Arabic)
darwish_tokens_nrc <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc_arabic, by = "word", relationship = "many-to-many")

# Tokenize and join to NRC -- Baraka (English)
baraka_tokens_nrc <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc_english, by = "word", relationship = "many-to-many")

# Report match rates
darwish_total <- darwish_poems %>% unnest_tokens(word, text) %>% nrow()
baraka_total  <- baraka_poems %>% unnest_tokens(word, text) %>% nrow()

cat("Darwish: ", nrow(darwish_tokens_nrc), "emotion matches from",
    darwish_total, "total tokens (",
    round(100 * n_distinct(darwish_tokens_nrc$word) /
            n_distinct(darwish_poems %>% unnest_tokens(word, text) %>% pull(word)), 1),
    "% unique words matched)\n")
cat("Baraka: ", nrow(baraka_tokens_nrc), "emotion matches from",
    baraka_total, "total tokens (",
    round(100 * n_distinct(baraka_tokens_nrc$word) /
            n_distinct(baraka_poems %>% unnest_tokens(word, text) %>% pull(word)), 1),
    "% unique words matched)\n")
```

```{r sentiment-aggregate}
# Aggregate emotion counts per poem
darwish_emotions <- darwish_tokens_nrc %>%
  count(poem, sentiment) %>%
  group_by(poem) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(poet = "Mahmoud Darwish")

baraka_emotions <- baraka_tokens_nrc %>%
  count(poem, sentiment) %>%
  group_by(poem) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(poet = "Amiri Baraka")

all_emotions <- bind_rows(darwish_emotions, baraka_emotions)
```

```{r sentiment-barplot, fig.height=7}
# Average emotion proportion by poet
emotion_summary <- all_emotions %>%
  group_by(poet, sentiment) %>%
  summarise(avg_prop = mean(proportion), .groups = "drop")

ggplot(emotion_summary,
       aes(x = reorder(sentiment, avg_prop), y = avg_prop, fill = poet)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Emotional Profile: Darwish vs. Baraka",
       subtitle = "Average proportion across all poems per poet",
       x = "Emotion / Sentiment", y = "Average Proportion",
       fill = "Poet") +
  theme_minimal(base_size = 13)
```

```{r sentiment-facet, fig.height=8}
# Only the 8 core emotions (exclude positive/negative)
core_emotions <- c("anger", "anticipation", "disgust", "fear",
                   "joy", "sadness", "surprise", "trust")

all_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  group_by(poet, sentiment) %>%
  summarise(avg_prop = mean(proportion),
            sd_prop  = sd(proportion), .groups = "drop") %>%
  ggplot(aes(x = sentiment, y = avg_prop, fill = poet)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = pmax(avg_prop - sd_prop, 0),
                    ymax = avg_prop + sd_prop),
                position = position_dodge(width = 0.8), width = 0.25) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Core Emotion Comparison (8 Emotions)",
       subtitle = "Mean +/- SD across poems",
       x = NULL, y = "Proportion", fill = "Poet") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Both poets share sadness and fear as their two dominant emotions, but the
distributions differ. Darwish's profile is led by sadness (13.6%) and fear (13.1%),
with surprise (11.9%) ranking third -- a pattern consistent with exile poetry,
where loss and uncertainty dominate. Baraka's top emotion is fear (12.6%), followed
by sadness (10.3%) and anger (10.2%); his anger score is notably close to his
sadness, reflecting a more confrontational register. Darwish's joy score (4.6%) is
roughly half of Baraka's (8.7%), suggesting a darker emotional baseline in the
Arabic corpus.


## 3.2 Lexical Diversity: TTR and MATTR

Type-Token Ratio (TTR) measures the proportion of unique words (types) relative to
total words (tokens) in a text. A higher TTR indicates more diverse vocabulary,
while a lower TTR suggests more repetition. Because TTR is a pure ratio computed
*within* each text, it is language-independent: a Darwish poem with TTR = 0.80 and
a Baraka poem with TTR = 0.65 are directly comparable.

Simple TTR is sensitive to text length (longer texts tend to have lower TTR), so we
also compute the Moving-Average TTR (MATTR), which uses a sliding window to provide
a more robust estimate across texts of varying length.

```{r ttr-compute}
# Build quanteda corpora with appropriate stopwords
darwish_corpus <- corpus(darwish_poems, text_field = "text",
                         docid_field = "poem")
baraka_corpus  <- corpus(baraka_poems, text_field = "text",
                         docid_field = "poem")

# Tokenize and remove stopwords (normalized for Arabic)
darwish_toks <- tokens(darwish_corpus, remove_punct = TRUE,
                       remove_numbers = TRUE) %>%
  tokens_remove(ar_stops) %>%
  tokens_select(min_nchar = 2)

baraka_toks <- tokens(baraka_corpus, remove_punct = TRUE,
                      remove_numbers = TRUE) %>%
  tokens_remove(en_stops) %>%
  tokens_select(min_nchar = 3)

# DFM
darwish_dfm <- dfm(darwish_toks)
baraka_dfm  <- dfm(baraka_toks)

# Lexical diversity — TTR from DFM, MATTR from tokens
darwish_ttr   <- textstat_lexdiv(darwish_dfm, measure = "TTR")
darwish_mattr <- textstat_lexdiv(darwish_toks, measure = "MATTR")
darwish_div   <- left_join(darwish_ttr, darwish_mattr, by = "document") %>%
  mutate(poet = "Mahmoud Darwish")

baraka_ttr   <- textstat_lexdiv(baraka_dfm, measure = "TTR")
baraka_mattr <- textstat_lexdiv(baraka_toks, measure = "MATTR")
baraka_div   <- left_join(baraka_ttr, baraka_mattr, by = "document") %>%
  mutate(poet = "Amiri Baraka")

all_div <- bind_rows(darwish_div, baraka_div)
```

```{r ttr-boxplot, fig.height=5}
# TTR boxplot
ggplot(all_div, aes(x = poet, y = TTR, fill = poet)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2.5) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Lexical Diversity: Type-Token Ratio (TTR)",
       subtitle = "Each point is one poem",
       y = "TTR (unique words / total words)", x = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

```{r mattr-boxplot, fig.height=5}
# MATTR boxplot
ggplot(all_div, aes(x = poet, y = MATTR, fill = poet)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2.5) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Lexical Diversity: Moving-Average TTR (MATTR)",
       subtitle = "More robust to poem length variation (window = 100 tokens)",
       y = "MATTR", x = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

```{r ttr-table}
# Summary statistics
div_summary <- all_div %>%
  group_by(poet) %>%
  summarise(
    mean_TTR   = round(mean(TTR, na.rm = TRUE), 3),
    median_TTR = round(median(TTR, na.rm = TRUE), 3),
    sd_TTR     = round(sd(TTR, na.rm = TRUE), 3),
    mean_MATTR   = round(mean(MATTR, na.rm = TRUE), 3),
    median_MATTR = round(median(MATTR, na.rm = TRUE), 3),
    sd_MATTR     = round(sd(MATTR, na.rm = TRUE), 3),
    .groups = "drop"
  )

kable(div_summary,
      col.names = c("Poet", "Mean TTR", "Median TTR", "SD TTR",
                     "Mean MATTR", "Median MATTR", "SD MATTR"),
      caption = "Lexical diversity summary by poet")
```

Darwish's mean TTR (0.805) slightly exceeds Baraka's (0.796), though Arabic
morphology -- which generates more distinct surface forms through prefixes and
suffixes -- likely inflates this figure. MATTR tells a slightly different story:
Baraka's mean MATTR (0.879) edges past Darwish's (0.854), suggesting comparable
diversity once text length is controlled for. The outliers are more telling than the
averages. Baraka's "SOS" has a TTR of just 0.27, driven by its heavy use of
onomatopoeia and repetitive incantation. Darwish's lowest-TTR poem is "Passers in
Passing Words" (0.63), where the deliberate anaphoric repetition of key phrases is
a rhetorical strategy, not a vocabulary limitation.


## 3.3 Sparsity and DFM Analysis

Sparsity measures the proportion of zeros in a Document-Feature Matrix (DFM). High
sparsity means most words appear in very few poems, indicating a vocabulary that is
spread thinly across the corpus. Lower sparsity means more shared vocabulary across
poems. This metric is computed within each poet's sub-corpus, making it comparable
across languages.

```{r sparsity-compute}
# Overall sparsity per poet
darwish_sparsity <- round(sparsity(darwish_dfm) * 100, 1)
baraka_sparsity  <- round(sparsity(baraka_dfm) * 100, 1)

# DFM dimensions
dfm_stats <- tibble(
  Poet = c("Mahmoud Darwish", "Amiri Baraka"),
  Documents = c(ndoc(darwish_dfm), ndoc(baraka_dfm)),
  Features  = c(nfeat(darwish_dfm), nfeat(baraka_dfm)),
  `Features per Doc` = round(c(nfeat(darwish_dfm) / ndoc(darwish_dfm),
                                nfeat(baraka_dfm) / ndoc(baraka_dfm))),
  `Sparsity (%)` = c(darwish_sparsity, baraka_sparsity)
)

kable(dfm_stats,
      col.names = c("Poet", "Documents", "Features", "Features per Doc", "Sparsity (%)"),
      caption = "Document-Feature Matrix comparison")
```

```{r sparsity-trim, fig.height=5}
# Sparsity at different trim levels
trim_levels <- seq(0, 0.5, by = 0.05)

sparsity_by_trim <- map_dfr(trim_levels, function(min_df) {
  d_trim <- dfm_trim(darwish_dfm, min_docfreq = max(min_df, 0.01),
                     docfreq_type = "prop")
  b_trim <- dfm_trim(baraka_dfm, min_docfreq = max(min_df, 0.01),
                     docfreq_type = "prop")
  tibble(
    min_docfreq = min_df,
    Darwish_sparsity = ifelse(nfeat(d_trim) > 0, sparsity(d_trim), NA_real_),
    Baraka_sparsity  = ifelse(nfeat(b_trim) > 0, sparsity(b_trim), NA_real_),
    Darwish_features = nfeat(d_trim),
    Baraka_features  = nfeat(b_trim)
  )
})

sparsity_by_trim %>%
  select(min_docfreq, Darwish_sparsity, Baraka_sparsity) %>%
  pivot_longer(-min_docfreq, names_to = "poet", values_to = "sparsity") %>%
  mutate(poet = str_remove(poet, "_sparsity")) %>%
  ggplot(aes(x = min_docfreq, y = sparsity, color = poet)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(labels = percent_format()) +
  labs(title = "Sparsity vs. Minimum Document Frequency Threshold",
       subtitle = "How sparsity changes as rare terms are progressively removed",
       x = "Minimum Document Frequency (proportion of poems)",
       y = "Sparsity", color = "Poet") +
  theme_minimal(base_size = 13)
```

```{r topfeatures}
# Top features per poet as formatted tables
darwish_top15 <- topfeatures(darwish_dfm, n = 15)
baraka_top15  <- topfeatures(baraka_dfm, n = 15)

kable(
  tibble(Word = names(darwish_top15), Count = unname(darwish_top15)),
  caption = "Top 15 features: Darwish (after stopword removal)"
)
kable(
  tibble(Word = names(baraka_top15), Count = unname(baraka_top15)),
  caption = "Top 15 features: Baraka (after stopword removal)"
)
```

Darwish's DFM sparsity (92.1%) is slightly higher than Baraka's (91.4%), meaning
that a given word in Darwish's corpus is less likely to appear across multiple poems.
Arabic morphology contributes to this gap: different inflected forms of the same root
are treated as separate features. The trim analysis shows how rapidly sparsity drops
when rare terms are removed, confirming that both corpora have a long tail of words
used in only one or two poems.


## 3.4 Word Frequency Analysis

Word frequency analysis counts the most common words after removing stopwords,
revealing the core thematic vocabulary of each poet. Since Arabic and English words
cannot be compared directly, we display each poet's top words separately in a
side-by-side visualization.

```{r word-freq}
# Darwish word frequencies (using normalized Arabic stopwords)
darwish_words <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% ar_stops,
         str_length(word) > 1) %>%
  count(word, sort = TRUE)

# Baraka word frequencies (using unified English stopwords)
baraka_words <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% en_stops,
         str_length(word) > 2) %>%
  count(word, sort = TRUE)
```

```{r freq-barplot, fig.height=7}
top_d <- darwish_words %>% slice_max(n, n = 15) %>%
  mutate(poet = "Mahmoud Darwish")
top_b <- baraka_words %>% slice_max(n, n = 15) %>%
  mutate(poet = "Amiri Baraka")

bind_rows(top_d, top_b) %>%
  ggplot(aes(x = reorder_within(word, n, poet), y = n, fill = poet)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poet, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Top 15 Most Frequent Words by Poet",
       subtitle = "After stopword removal",
       x = NULL, y = "Count") +
  theme_minimal(base_size = 13)
```

```{r wordclouds, fig.height=10, fig.width=16}
wc_data <- bind_rows(
  darwish_words %>% slice_max(n, n = 80) %>% mutate(poet = "Mahmoud Darwish"),
  baraka_words  %>% slice_max(n, n = 80) %>% mutate(poet = "Amiri Baraka")
)

ggplot(wc_data, aes(label = word, size = n, color = n)) +
  geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle") +
  scale_size_area(max_size = 24) +
  scale_color_gradient(low = "darkgreen", high = "darkred") +
  facet_wrap(~poet, scales = "free") +
  theme_minimal(base_size = 14) +
  labs(title = "Word Clouds: Darwish vs. Baraka",
       subtitle = "Top 80 words after stopword removal")
```

Darwish's most frequent content words after stopword removal are "look out" (اطل),
"and I" (وانا), "my father" (ابي), "the sea" (البحر), and "and you" (وانت) -- a
vocabulary rooted in the self, landscape, family, and address. Baraka's top words -- "black," "people,"
"killed," "poems," "world" -- reflect a vocabulary oriented toward collective
identity and violence. The contrast is stark: Darwish's lexicon is intimate and
geographic, while Baraka's is communal and confrontational.


## 3.5 Bigram Network Graph

Bigram analysis extracts consecutive word pairs and visualizes them as network
graphs, where nodes represent words and edges connect words that appear together.
This reveals phrase structure, recurring collocations, and rhetorical patterns.
As with word frequency, bigrams are language-specific and must be analyzed
separately per poet.

```{r bigrams}
# Darwish bigrams (normalized Arabic stopwords)
darwish_bigrams <- darwish_poems %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% ar_stops, !word2 %in% ar_stops,
         str_length(word1) > 1, str_length(word2) > 1) %>%
  count(word1, word2, sort = TRUE)

# Baraka bigrams (unified English stopwords)
baraka_bigrams <- baraka_poems %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% en_stops, !word2 %in% en_stops,
         str_length(word1) > 2, str_length(word2) > 2) %>%
  count(word1, word2, sort = TRUE)

cat("Darwish unique bigrams:", nrow(darwish_bigrams), "\n")
cat("Baraka unique bigrams:", nrow(baraka_bigrams), "\n")
```

```{r bigram-table}
kable(head(darwish_bigrams, 10),
      col.names = c("Word 1", "Word 2", "Count"),
      caption = "Top 10 Darwish bigrams (after stopword removal)")
kable(head(baraka_bigrams, 10),
      col.names = c("Word 1", "Word 2", "Count"),
      caption = "Top 10 Baraka bigrams (after stopword removal)")
```

```{r bigram-network, fig.height=8, fig.width=14}
# Network plotting function
plot_bigram_network <- function(bigram_counts, title_text,
                                 n_edges = 25, min_n = 2) {
  graph_data <- bigram_counts %>%
    filter(n >= min_n) %>%
    slice_max(n, n = n_edges)

  if (nrow(graph_data) == 0) {
    # If no bigrams meet threshold, lower it
    graph_data <- bigram_counts %>% slice_max(n, n = n_edges)
  }

  graph <- graph_from_data_frame(graph_data)

  ggraph(graph, layout = "fr") +
    geom_edge_link(
      aes(edge_alpha = n, edge_width = n),
      show.legend = FALSE,
      arrow = arrow(type = "closed", length = unit(0.12, "inches")),
      end_cap = circle(3, "mm")
    ) +
    geom_node_point(color = "steelblue", size = 5, alpha = 0.8) +
    geom_node_text(aes(label = name), vjust = 1.2, hjust = 0.5,
                   size = 3.5, check_overlap = TRUE) +
    theme_void(base_size = 13) +
    labs(title = title_text)
}

# Side-by-side networks
set.seed(42)
p_darwish <- plot_bigram_network(darwish_bigrams,
                                  "Darwish: Bigram Network",
                                  n_edges = 30, min_n = 2)
p_baraka  <- plot_bigram_network(baraka_bigrams,
                                  "Baraka: Bigram Network",
                                  n_edges = 30, min_n = 2)

p_darwish + p_baraka +
  plot_annotation(
    title = "Bigram Network Graphs: Darwish vs. Baraka",
    subtitle = "Nodes = words, edges = consecutive word pairs (after stopword removal)",
    theme = theme(plot.title = element_text(size = 16, face = "bold"))
  )
```

Baraka's top bigram is "black people," which functions as both subject and rallying
cry across multiple poems. His network also shows the striking pair "survive
blackman / blackman survive" -- a chiastic repetition typical of his incantatory
style. Darwish's top bigrams include "think of others" (فكر بغيرك, also a poem
title used as a refrain) and "passing words" (الكلمات العابره). Where Baraka's
network clusters tightly around racial identity, Darwish's spreads across
landscape and temporal imagery.


## 3.6 Radar Chart: Emotion Profile Comparison

A radar (spider) chart places both poets' average emotion proportions on the same
radial axes, making it immediately visible where one poet dominates and where they
converge. This is the single most informative cross-linguistic visualization in our
analysis.

```{r radar-chart, fig.height=7, fig.width=8}
# Prepare data in fmsb format: one row per poet, columns = emotions
radar_data <- all_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  group_by(poet, sentiment) %>%
  summarise(avg = mean(proportion), .groups = "drop") %>%
  pivot_wider(names_from = sentiment, values_from = avg)

# fmsb requires: row1 = max, row2 = min, then data rows
radar_max <- rep(max(unlist(radar_data[, -1])) * 1.15, 8)
radar_min <- rep(0, 8)

radar_df <- rbind(radar_max, radar_min, radar_data[, -1])
rownames(radar_df) <- c("max", "min", radar_data$poet)

# Plot
colors_fill <- c(rgb(0.8, 0.2, 0.2, 0.25), rgb(0.2, 0.4, 0.8, 0.25))
colors_line <- c(rgb(0.8, 0.2, 0.2, 0.9),  rgb(0.2, 0.4, 0.8, 0.9))

radar_max_val <- max(unlist(radar_data[, -1])) * 1.15
radarchart(radar_df,
           axistype = 1,
           pcol = colors_line, pfcol = colors_fill, plwd = 2.5, plty = 1,
           cglcol = "grey70", cglty = 1, cglwd = 0.8,
           axislabcol = "grey40",
           caxislabels = paste0(round(seq(0, radar_max_val, length.out = 5) * 100, 1), "%"),
           vlcex = 1.1,
           title = "Emotion Radar: Darwish vs. Baraka")
legend("topright",
       legend = radar_data$poet,
       col = colors_line, lwd = 2.5, lty = 1,
       fill = colors_fill, border = colors_line,
       bty = "n", cex = 1.0)
```

The two polygons overlap substantially on fear and sadness -- the shared emotional
core of resistance poetry. The largest divergence appears on the surprise axis, where
Darwish's polygon extends well beyond Baraka's (11.9% vs. 6.8%), and on joy, where
Baraka reaches nearly twice Darwish's score. The radar makes visible what the bar
charts only imply: Darwish writes from a darker, more uncertain emotional place,
while Baraka balances his anger with more moments of affirmation.


## 3.7 Poem-Emotion Heatmap

The heatmap shows the emotion proportion for each individual poem, revealing which
specific poems drive the aggregate patterns. This identifies emotionally distinctive
poems and shows within-poet variation.

```{r heatmap-darwish, fig.height=7, fig.width=10}
# Darwish heatmap
darwish_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  mutate(poem = str_trunc(poem, 30)) %>%
  ggplot(aes(x = sentiment, y = reorder(poem, proportion), fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(proportion, 2)), size = 3, color = "grey20") +
  scale_fill_gradient(low = "white", high = "#B22222",
                      labels = percent_format()) +
  labs(title = "Darwish: Emotion Heatmap by Poem",
       subtitle = "Proportion of emotion words in each poem",
       x = NULL, y = NULL, fill = "Proportion") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.grid = element_blank())
```

```{r heatmap-baraka, fig.height=7, fig.width=10}
# Baraka heatmap
baraka_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  mutate(poem = str_trunc(poem, 30)) %>%
  ggplot(aes(x = sentiment, y = reorder(poem, proportion), fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(proportion, 2)), size = 3, color = "grey20") +
  scale_fill_gradient(low = "white", high = "#1565C0",
                      labels = percent_format()) +
  labs(title = "Baraka: Emotion Heatmap by Poem",
       subtitle = "Proportion of emotion words in each poem",
       x = NULL, y = NULL, fill = "Proportion") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.grid = element_blank())
```

The heatmaps show that the corpus-level averages mask wide variation within each
poet. In Darwish's heatmap, "I Am Joseph, O Father" (أنا يوسف يا أبي) is dominated
by surprise (0.50), while "Rita and the Rifle" (ريتا والبندقية) peaks on sadness (0.33)
and anger (0.17). In Baraka's heatmap, "Black Art" stands out for its high fear (0.16)
and anger (0.13), while "Preface to a Twenty Volume Suicide Note" is highest on
trust (0.21) -- an unexpected result for a poem with "suicide" in the title. This
poem-level granularity matters: it shows that labeling either poet with a single
emotional tag misrepresents the range of their work.


## 3.8 TF-IDF: Distinctive Words

Term Frequency-Inverse Document Frequency (TF-IDF) identifies words that are
distinctive to each poet rather than just frequent. A word with high TF-IDF appears
often in one poet's work but rarely (or never) in the other's. This is computed by
treating each poet's full corpus as a single document. Since Arabic and English
never share words, TF-IDF is applied within each language separately across poems.

```{r tfidf, fig.height=8, fig.width=12}
# TF-IDF per poem within each poet
darwish_tfidf <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% ar_stops, str_length(word) > 1) %>%
  count(poem, word, sort = TRUE) %>%
  bind_tf_idf(word, poem, n)

baraka_tfidf <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% en_stops, str_length(word) > 2) %>%
  count(poem, word, sort = TRUE) %>%
  bind_tf_idf(word, poem, n)

# Top TF-IDF words per poet (across all poems)
top_tfidf_d <- darwish_tfidf %>%
  arrange(desc(tf_idf)) %>%
  distinct(word, .keep_all = TRUE) %>%
  slice_max(tf_idf, n = 15) %>%
  mutate(poet = "Mahmoud Darwish")

top_tfidf_b <- baraka_tfidf %>%
  arrange(desc(tf_idf)) %>%
  distinct(word, .keep_all = TRUE) %>%
  slice_max(tf_idf, n = 15) %>%
  mutate(poet = "Amiri Baraka")

bind_rows(top_tfidf_d, top_tfidf_b) %>%
  ggplot(aes(x = reorder_within(word, tf_idf, poet),
             y = tf_idf, fill = poet)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poet, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Most Distinctive Words by Poet (TF-IDF)",
       subtitle = "Words that are uniquely important to individual poems",
       x = NULL, y = "TF-IDF Score") +
  theme_minimal(base_size = 13)
```

```{r tfidf-per-poem, fig.height=9, fig.width=13}
# Top TF-IDF words per poem -- Baraka (4 poems)
baraka_tfidf %>%
  filter(poem %in% c("Somebody Blew Up America", "Black Art",
                      "Preface to a Twenty Volume Suicide Note",
                      "An Agony. As Now.")) %>%
  group_by(poem) %>%
  slice_max(tf_idf, n = 8) %>%
  ungroup() %>%
  ggplot(aes(x = reorder_within(word, tf_idf, poem),
             y = tf_idf, fill = poem)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poem, scales = "free", ncol = 2) +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Baraka: Most Distinctive Words per Poem (TF-IDF)",
       subtitle = "Words that define each poem's unique vocabulary",
       x = NULL, y = "TF-IDF") +
  theme_minimal(base_size = 12)
```

Where frequency analysis shows what is common across a poet's work, TF-IDF shows
what makes each poem unique. The per-poem TF-IDF plots for Baraka confirm that
"Somebody Blew Up America" is dominated by accusatory vocabulary, while "Preface"
is defined by more introspective terms. These thematic signatures help distinguish
poems that share high-frequency words but differ in subject and tone.


## 3.9 Sentiment Arc: Emotional Trajectory

The sentiment arc tracks how positive and negative sentiment shifts across the lines
of a poem, revealing its emotional structure. We compute a rolling sentiment score
by dividing each poem into segments and calculating the net sentiment
(positive minus negative) in each segment.

```{r sentiment-arc, fig.height=8, fig.width=12}
# Helper: compute sentiment arc for a set of poems + NRC lexicon
compute_arc <- function(poems_df, nrc_lex, poet_name, n_segments = 10) {
  poems_df %>%
    unnest_tokens(word, text) %>%
    group_by(poem) %>%
    mutate(word_pos = row_number(),
           total_words = n(),
           segment = ceiling(word_pos / total_words * n_segments)) %>%
    ungroup() %>%
    inner_join(nrc_lex %>% filter(sentiment %in% c("positive", "negative")),
               by = "word", relationship = "many-to-many") %>%
    mutate(score = ifelse(sentiment == "positive", 1, -1)) %>%
    group_by(poem, segment) %>%
    summarise(net_sentiment = sum(score), .groups = "drop") %>%
    mutate(poet = poet_name)
}

darwish_arc <- compute_arc(darwish_poems, nrc_arabic,
                            "Mahmoud Darwish", n_segments = 10)
baraka_arc  <- compute_arc(baraka_poems, nrc_english,
                            "Amiri Baraka", n_segments = 10)

# Select a few representative poems per poet
darwish_show <- c("بطاقة هوية", "عاشق من فلسطين",
                  "عابرون في كلام عابر", "فكر بغيرك")
baraka_show  <- c("Somebody Blew Up America", "Black Art",
                  "Preface to a Twenty Volume Suicide Note",
                  "An Agony. As Now.")

bind_rows(
  darwish_arc %>% filter(poem %in% darwish_show),
  baraka_arc  %>% filter(poem %in% baraka_show)
) %>%
  ggplot(aes(x = segment, y = net_sentiment, color = poem)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  facet_wrap(~poet, scales = "free_y", ncol = 1) +
  scale_x_continuous(breaks = 1:10,
                     labels = paste0(seq(10, 100, 10), "%")) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Sentiment Arc: Emotional Trajectory Across Poems",
       subtitle = "Net sentiment (positive - negative) across 10 equal segments of each poem",
       x = "Position in Poem", y = "Net Sentiment",
       color = "Poem") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 10))
```

The arcs add a temporal dimension invisible in aggregate counts. Darwish's
"Identity Card" (بطاقة هوية) oscillates between positive and negative segments,
reflecting its alternating structure of defiance and lament. Baraka's "Black Art"
stays below the zero line for most of its length (segments 1--8 are all negative)
before spiking sharply positive at the end -- a dramatic reversal that mirrors the
poem's shift from denouncing to envisioning. These arc shapes correspond to
different rhetorical strategies that aggregate emotion counts cannot capture.


## 3.10 Pairwise Word Correlation

Pairwise correlation measures which words tend to co-occur within the same poems.
Unlike bigrams (which capture adjacent words), correlation captures words that
appear in the same poems regardless of proximity. We use the `widyr` package's
`pairwise_cor()` function with a phi coefficient, filtering for words that appear
in at least 3 poems.

```{r pairwise-cor, fig.height=8, fig.width=12}
# Baraka pairwise correlation (English -- more interpretable for display)
baraka_word_pairs <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% en_stops, str_length(word) > 2) %>%
  add_count(word) %>%
  filter(n >= 5) %>%
  pairwise_cor(word, poem, sort = TRUE)

# Top positive correlations
set.seed(42)
baraka_word_pairs %>%
  filter(correlation > 0.4) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation, edge_width = correlation),
                 show.legend = TRUE, color = "steelblue") +
  geom_node_point(size = 5, color = "#E74C3C", alpha = 0.8) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3.5,
                 max.overlaps = 15) +
  scale_edge_width(range = c(0.5, 2.5)) +
  theme_void(base_size = 13) +
  labs(title = "Baraka: Word Correlation Network",
       subtitle = "Words that co-occur in the same poems (phi > 0.4)",
       edge_alpha = "Correlation", edge_width = "Correlation")
```

```{r pairwise-darwish, fig.height=8, fig.width=12}
# Darwish pairwise correlation
darwish_word_pairs <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% ar_stops, str_length(word) > 1) %>%
  add_count(word) %>%
  filter(n >= 5) %>%
  pairwise_cor(word, poem, sort = TRUE)

set.seed(42)
darwish_word_pairs %>%
  filter(correlation > 0.4) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation, edge_width = correlation),
                 show.legend = TRUE, color = "darkgreen") +
  geom_node_point(size = 5, color = "#2980B9", alpha = 0.8) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3.5,
                 max.overlaps = 15) +
  scale_edge_width(range = c(0.5, 2.5)) +
  theme_void(base_size = 13) +
  labs(title = "Darwish: Word Correlation Network",
       subtitle = "Words that co-occur in the same poems (phi > 0.4)",
       edge_alpha = "Correlation", edge_width = "Correlation")
```

Unlike bigrams, which capture adjacency, pairwise correlation captures words that
co-occur in the same poems even when separated by many lines. In Baraka's network,
words related to racial identity (e.g., "black," "people," "white") cluster tightly,
while a separate cluster forms around political accusation vocabulary. Darwish's
network tends to group landscape and body imagery together. These clusters map the
thematic modules of each poet's work at a level that individual word counts cannot
reach.


# 4. Discussion

## 4.1 Emotional Registers of Resistance

The NRC emotion analysis is the most direct cross-linguistic comparison in this
study. Both poets share fear and sadness as their dominant emotions, confirming that
resistance poetry -- whether rooted in Palestinian exile or African-American
struggle -- draws heavily on these registers. The difference lies in the secondary
emotions: Darwish leans toward surprise (11.9%), reflecting the uncertainty and
displacement of exile, while Baraka's anger (10.2%) runs nearly as strong as his
sadness, consistent with his confrontational poetics.

One caveat: not every word matches the NRC lexicon. Arabic's morphological
complexity reduces match rates because prefixes and suffixes create surface forms
absent from the lexicon. We report match rates transparently so the reader can
judge the reliability of these profiles.

## 4.2 Vocabulary and Structure

Mean TTR values are close (Darwish 0.805, Baraka 0.796), but the outliers tell a
richer story. Baraka's "SOS" has a TTR of just 0.27 -- the lowest in either
corpus -- because its heavy onomatopoeia and incantatory repetition deliberately
sacrifice vocabulary breadth for rhythmic force. Darwish uses
anaphora similarly in "Passers in Passing Words" (TTR 0.63), where the repeated
address structure is a rhetorical choice, not a vocabulary limitation. In both cases,
low TTR flags deliberate poetic strategy.

Sparsity is comparable across the two sub-corpora (Darwish 92.1%, Baraka 91.4%),
indicating that both poets draw on largely poem-specific vocabulary rather than
recycling a fixed set of terms. This high sparsity is expected in poetry, where each
poem typically addresses a distinct theme or occasion.

## 4.3 Thematic Vocabulary and Phrase Patterns

The word frequency and bigram analyses run separately per language but
expose a clear thematic contrast. Darwish's top words (اطل "look out," ابي
"my father," البحر "the sea") are intimate and landscape-oriented. Baraka's
("black," "people," "killed") are collective and political. The bigram "black
people" dominates Baraka's network as both subject and rallying cry, while
Darwish's top bigrams ("think of others," "passing words") are more reflective.

Word frequency and bigrams complement each other: frequency shows what each poet
returns to, while bigrams show how those terms combine into phrases. Baraka's
chiastic pair "survive blackman / blackman survive" is invisible in a word count
but central to his rhythmic strategy.

## 4.4 Advanced Cross-Linguistic Patterns

The radar chart makes the shape difference between the two poets immediately
visible. Both polygons extend furthest on sadness and fear, but Darwish's polygon
bulges outward on surprise while Baraka's extends on trust and joy. This shape
contrast -- Darwish darker and more uncertain, Baraka more balanced between anger
and affirmation -- is harder to see in side-by-side bar charts.

The poem-level heatmaps show that these aggregate profiles mask wide internal
variation. Darwish's "I Am Joseph, O Father" is dominated by surprise (0.50), while
"Rita and the Rifle" peaks on sadness and anger. Baraka's "Black Art" concentrates
on fear and anger, while "Preface to a Twenty Volume Suicide Note" scores highest
on trust (0.21). Neither poet is a one-note writer, even though both are often filed
under "resistance."

TF-IDF and the per-poem breakdown shift attention from shared
vocabulary to distinctive vocabulary. Words that define "Somebody Blew Up America"
differ sharply from those that define "Preface," even though both poems share many
high-frequency terms. This technique helps separate thematic identity from mere
frequency.

The sentiment arcs capture emotional trajectory -- something invisible in aggregate
counts. Baraka's "Black Art" sustains a negative arc for eight segments before
spiking positive at the end, while Darwish's "Identity Card" oscillates throughout.
These shape differences reflect distinct rhetorical strategies: sustained buildup
versus alternating tension.

Pairwise correlation networks group words that share thematic space across poems,
regardless of adjacency. Baraka's network clusters racial identity vocabulary
separately from political accusation vocabulary. Darwish's clusters landscape
imagery with body imagery. These groupings map the latent thematic modules that
organize each poet's work.

## 4.5 Limitations

We note several methodological limitations:

1. **Arabic morphology**: Arabic's rich inflectional system produces many surface
   forms from a single root. Without lemmatization or stemming (which would require
   specialized Arabic NLP tools beyond the scope of this course), each inflected form
   is treated as a separate word. This may inflate TTR and sparsity for Darwish
   relative to Baraka, and reduce NRC lexicon match rates.

2. **Corpus size**: With 14 poems per poet, this is a small corpus by text mining
   standards. Results are exploratory and should not be generalized to the full body
   of either poet's work.

3. **NRC lexicon coverage**: The multilingual NRC lexicon relies on machine
   translation for non-English entries. Some Arabic emotion assignments may be
   imprecise due to translation artifacts or cultural differences in emotional
   connotation.

4. **Translation equivalence**: We compare numeric metrics, not meanings. Two poems
   might score identically on "sadness" for entirely different reasons, rooted in
   different cultural contexts of loss.


# 5. Conclusion

Darwish and Baraka share sadness and fear as their dominant emotions, but diverge
on secondary registers: Darwish's work carries more surprise and less joy, while
Baraka balances anger with trust and affirmation. Their lexical strategies also
differ -- Darwish's vocabulary is intimate and geographic, Baraka's is collective
and confrontational -- though both poets show high within-corpus variation, with
individual poems occupying very different emotional and thematic positions.

The ten techniques applied here -- NRC sentiment, radar charts, heatmaps, sentiment
arcs, TF-IDF, TTR/MATTR, sparsity, word frequency, bigrams, and pairwise
correlation -- span tokenization, sentiment analysis, lexical diversity,
document-feature matrices, and network analysis. The NRC multilingual lexicon
proved essential for bridging the language divide, enabling direct comparison
of emotional profiles between Arabic and English poetry.

The poem-level analyses (heatmaps, sentiment arcs, TF-IDF) are where the most
interesting patterns emerge: they show that neither poet is monolithic, and that
corpus-level averages obscure substantial internal variation. These quantitative
patterns complement traditional literary analysis by making visible the emotional
and structural architecture of resistance poetry across two very different
linguistic traditions.


# References

- Benoit, K. et al. (2018). quanteda: An R package for the quantitative analysis
  of textual data. *Journal of Open Source Software*, 3(30), 774.
- Covington, M. A. & McFall, J. D. (2010). Cutting the Gordian Knot: The
  Moving-Average Type-Token Ratio (MATTR). *Journal of Quantitative Linguistics*,
  17(2), 94-100.
- Mohammad, S. M. & Turney, P. D. (2013). Crowdsourcing a Word-Emotion Association
  Lexicon. *Computational Intelligence*, 29(3), 436-465.
- Reagan, A. J. et al. (2016). The emotional arcs of stories are dominated by six
  basic shapes. *EPJ Data Science*, 5(1), 1-12.
- Silge, J. & Robinson, D. (2017). *Text Mining with R: A Tidy Approach*. O'Reilly.
- Welbers, K. et al. (2017). Text Analysis in R. *Communication Methods and Measures*,
  11(4), 245-265.
- Poem sources: aldiwan.net, poetsgate.com, poetryfoundation.org, poemhunter.com,
  Genius, becauseofthemwecan.com

---

*Analysis conducted with R `r getRversion()`, using tidytext, quanteda, and the
NRC Emotion Lexicon.*

```{r session-info}
sessionInfo()
```
