---
title: "Voices of Resistance: A Comparative Text Analysis of Amiri Baraka and Mahmoud Darwish"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(quanteda)
library(tidytext)
library(arabicStemR)  # for Arabic morphological processing
```

---

## 1. Introduction

This analysis undertakes a computational comparison of six poems — three by the Palestinian poet **Mahmoud Darwish** (in Arabic) and three by the African-American poet **Amiri Baraka** (in English). Both poets are canonical voices of political resistance, but they operate in different languages, traditions, and geopolitical contexts. The central question driving this analysis is:

> **Does the linguistic and structural nature of Arabic versus English shape how anger, defiance, and hope are expressed in resistance poetry?**

We approach this question through two primary analytical lenses: **lexical diversity** and **sentiment trajectory**.

### 1.2 downloading NRC arabic library

```{r}
# Step 2: Load it as a custom lexicon
nrc_arabic <- read_tsv("NRC-Emotion-Lexicon/NRC-Emotion-Lexicon-ForVariousLanguages.txt") %>%
  select(word = `Arabic`,
         anger, anticipation, disgust, fear,
         joy, negative, positive, sadness, surprise, trust) %>%
  pivot_longer(
    cols = -word,
    names_to = "sentiment",
    values_to = "value"
  ) %>%
  filter(value == 1) %>%
  select(word, sentiment)
```
---

## 2. Data


### 2.1 Baraka Corpus

The three Baraka poems included in this analysis are:

- *SOS*
- *Ka'Ba*
- *Somebody Blew Up America*

These were compiled into a tidy line-level dataset using the `baraka_poems_dataset.R` script. Each row represents one line, tagged with poet, poem title, and line number.

```{r}
baraka_poems <- source("~/Documents/GitHub/text_mining_uc3m/Baraka_dataset.R")
```


### 2.2 Darwish Corpus

The three Darwish poems are:

- [Poem 1 title]
- [Poem 2 title]
- [Poem 3 title]

**Note on Arabic preprocessing:** Arabic text requires special handling before analysis. The following steps must be applied before building the corpus:

- **Normalization:** Standardize Alef variants (أ إ آ → ا), remove diacritics (tashkeel), and normalize Ta Marbuta.
- **Tokenization:** Use a whitespace-aware tokenizer that respects Arabic right-to-left script.
- **Stopword removal:** Apply an Arabic stopword list (e.g., from the `stopwords` package with `language = "ar"`).
- **Stemming/Lemmatization:** Consider whether to stem (reduces words to root) or lemmatize (reduces to dictionary form). Stemming will increase TTR artificially — document your choice.

```{r}
source("~/Documents/GitHub/text_mining_uc3m/darwish_dataset.R")
```


### 2.3 Combined Dataset

```{r data-load}
# Load Baraka line-level data
# source("baraka_poems_dataset.R")

# Load Darwish data
# darwish_lines <- read_csv("darwish_poems_dataset.csv")

# Combine
# all_lines <- bind_rows(baraka_lines, darwish_lines)
```

---

## 3. Analytical Framework

This project uses a **hybrid Quanteda + tidytext workflow**:

- **Quanteda** is used as the computational engine to build Document-Feature Matrices (DFMs) and calculate statistics that require matrix-level operations (e.g., TTR, document similarity).
- **tidytext** is used to reshape data into long/tidy format for visualization and sentiment mapping.

The two techniques are complementary: Quanteda handles the math correctly for Arabic, while tidytext integrates cleanly with `ggplot2` for output.

---

## 4. Technique 1 — Lexical Diversity (TTR)

### 4.1 What We Are Measuring

The **Type-Token Ratio (TTR)** measures the proportion of unique words (types) to total words (tokens) in a text:

$$TTR = \frac{\text{Number of Unique Words}}{\text{Total Number of Words}}$$

A higher TTR indicates a richer, more varied vocabulary. We expect Arabic to yield a higher TTR partly due to its morphological richness — a single Arabic root can generate many surface forms that count as distinct types.

**Important caveat:** TTR is sensitive to text length. Longer texts almost always have lower TTR. We will use **Moving Average TTR (MATTR)** or **Measure of Textual Lexical Diversity (MTLD)** from Quanteda to correct for this, enabling a fair comparison across poems of different lengths.

### 4.2 Building the DFM

```{r dfm-build}
# Step 1: Create a quanteda corpus from the combined dataset
# corpus_all <- corpus(all_lines, text_field = "line",
#                      docvars = data.frame(poet = all_lines$poet,
#                                           poem = all_lines$poem))

# Step 2: Tokenize (handling Arabic RTL correctly)
# tokens_all <- tokens(corpus_all,
#                      remove_punct = TRUE,
#                      remove_numbers = TRUE)

# Step 3: Remove stopwords (both languages)
# tokens_all <- tokens_remove(tokens_all,
#                              pattern = c(stopwords("en"), stopwords("ar")))

# Step 4: Build DFM
# dfm_all <- dfm(tokens_all)
```

### 4.3 Calculating TTR per Poem

```{r ttr-calculate}
# Group by poem before calculating
# dfm_by_poem <- dfm_group(dfm_all, groups = interaction(poet, poem))

# Calculate TTR
# ttr_results <- textstat_lexdiv(dfm_by_poem, measure = "TTR")
# mtld_results <- textstat_lexdiv(dfm_by_poem, measure = "MTLD")
```

### 4.4 Visualization

```{r ttr-plot}
# ttr_results %>%
#   mutate(poet = if_else(str_detect(document, "Baraka"), "Baraka", "Darwish")) %>%
#   ggplot(aes(x = reorder(document, TTR), y = TTR, fill = poet)) +
#   geom_col() +
#   coord_flip() +
#   labs(title = "Lexical Diversity (TTR) by Poem",
#        subtitle = "Higher = more varied vocabulary",
#        x = NULL, y = "Type-Token Ratio") +
#   theme_minimal()
```

**Interpretation guide:** When writing up results, address:

- Which poet has a higher average TTR, and is the difference consistent across poems?
- Does poem length explain any differences? (Check with MTLD which controls for length.)
- What does this suggest about Arabic morphology vs. English in resistance poetry?

---

## 5. Technique 2 — Sentiment Trajectory & Polarity

### 5.1 What We Are Measuring

Rather than a single sentiment score per poem, we map **how sentiment evolves across the poem** — from the first line to the last. This allows us to answer:

- Does Darwish build toward a peak of anger, or sustain it throughout?
- Does Baraka's anger follow a different structural arc?
- Where does *hope* or *defiance* appear relative to *grief* or *rage*?

We will track three dimensions of sentiment: **Anger**, **Hope/Defiance**, and **Grief/Sorrow**.

### 5.2 Lexicon Strategy

For **English (Baraka):** Use the NRC Emotion Lexicon (`tidytext::get_sentiments("nrc")`), which provides word-level scores for 8 emotions including anger, fear, trust, and anticipation.

For **Arabic (Darwish):** This requires a dedicated Arabic sentiment lexicon. Options include:

- **Arabic NRC Lexicon** (Mohammad et al.) — a translated version of NRC
- **ASTD** (Arabic Sentiment Tweets Dataset)
- **ArSentD-LEV** — if poems use Levantine dialect

Document which lexicon is used and note any coverage gaps (words in the poem not found in the lexicon).

### 5.3 Tidytext Sentiment Join

```{r sentiment-join}
# English: join NRC lexicon
# baraka_sentiment <- baraka_lines %>%
#   unnest_tokens(word, line) %>%
#   inner_join(get_sentiments("nrc"), by = "word") %>%
#   filter(sentiment %in% c("anger", "anticipation", "trust", "sadness")) %>%
#   group_by(poem, line_number, sentiment) %>%
#   summarise(n = n(), .groups = "drop")

# Arabic: join Arabic NRC or chosen lexicon
# darwish_sentiment <- darwish_lines %>%
#   unnest_tokens(word, line) %>%
#   inner_join(arabic_nrc, by = "word") %>%
#   ...
```

### 5.4 Smoothed Trajectory Plot

```{r sentiment-trajectory}
# all_sentiment %>%
#   filter(sentiment == "anger") %>%
#   ggplot(aes(x = line_number, y = n, color = poet)) +
#   geom_smooth(method = "loess", se = TRUE) +
#   facet_wrap(~poem, scales = "free_x") +
#   labs(title = "Anger Trajectory Across Poem Lines",
#        x = "Line Number", y = "Anger Word Count") +
#   theme_minimal()
```

**Note on smoothing:** Use `geom_smooth(method = "loess")` to reveal the overall arc without overreacting to individual lines. Show the raw points as well with `geom_point(alpha = 0.3)` for transparency.

### 5.5 Polarity Heatmap (Optional Extension)

A heatmap showing each poem as a column and each line as a row, colored by dominant emotion, can give an immediate visual intuition of structural differences between the poets.

```{r heatmap}
# all_sentiment %>%
#   group_by(poet, poem, line_number) %>%
#   slice_max(n, n = 1) %>%  # dominant emotion per line
#   ggplot(aes(x = poem, y = line_number, fill = sentiment)) +
#   geom_tile() +
#   scale_fill_brewer(palette = "Set2") +
#   facet_wrap(~poet, scales = "free") +
#   labs(title = "Dominant Emotion by Line",
#        x = NULL, y = "Line Number") +
#   theme_minimal()
```

---

## 6. Synthesis & Interpretation

This section should bring the two techniques into dialogue. Suggested questions to address:

- **Do the TTR results and sentiment results tell a consistent story?** For example, if Darwish has higher lexical diversity AND a more gradual emotional build, this might suggest a more layered rhetorical strategy.
- **What are the limitations?** Sentiment lexicons are imperfect; translation of emotion concepts across cultures is contested. Acknowledge that "anger" in the NRC lexicon was built on English-language intuitions.
- **What does "manifestation of anger" mean differently for each poet?** Consider the poems' political contexts — Darwish writing under occupation, Baraka writing in the aftermath of 9/11.

---

## 7. Next Steps & Open Questions

- [ ] Finalize Darwish poem selection and build `darwish_lines` dataset in the same structure as `baraka_lines`
- [ ] Evaluate Arabic lexicon options and document coverage rates
- [ ] Decide on TTR correction method (MATTR vs. MTLD) and justify in write-up
- [ ] Run pilot TTR calculation on Baraka alone to verify the Quanteda pipeline before adding Arabic
- [ ] Determine whether to stem Arabic text and document effect on TTR
- [ ] Draft interpretation of results for Section 6 once analysis is complete

---

## Appendix: Session Info

```{r session-info}
sessionInfo()
```
