---
title: "The Poetry of Resistance: A Cross-Linguistic Text Mining Analysis"
subtitle: "Mahmoud Darwish (Arabic) vs. Amiri Baraka (English)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: flatly
    code_folding: hide
    css: style.css
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 6, fig.align = "center",
  dev = if (requireNamespace("ragg", quietly = TRUE)) "ragg_png" else "png",
  dpi = 150
)
```

:::::::::::::::::: project-header
:::: team-card
<h4>Team</h4>

<ul>

<li>[Abdullah Tadmuri]{.member-name}</li>

<li>[Maiheliya Maimaitimin]{.member-name}</li>

<li>[Tone Varberg Sabri]{.member-name}</li>

</ul>

::: course-line
Text Mining Final Project · UC3M · 2025–2026
:::
::::

::::::::::::::: stats-grid
::::: stat-card
::: stat-number
28
:::

::: stat-label
Poems Analyzed
:::
:::::

::::: stat-card
::: stat-number
6
:::

::: stat-label
Techniques Used
:::
:::::

::::: stat-card
::: stat-number
2
:::

::: stat-label
Languages
:::
:::::

::::: stat-card
::: stat-number
14
:::

::: stat-label
Poems per Poet
:::
:::::
:::::::::::::::
::::::::::::::::::

# 1. Introduction

This study analyzes two prominent resistance poets from distinct
traditions: **Mahmoud Darwish** (1941--2008), the Palestinian national
poet who wrote in Arabic, and **Amiri Baraka** (1934--2014), the
African-American poet and activist who wrote in English.

Both poets use verse as a tool for collective resistance, though they
address different, but comparable struggles. Darwish's poetry centers on
Palestinian dispossession, exile, and the longing for homeland, while
Baraka's work confronts anti-Black racism, celebrates Black identity,
and calls for revolutionary change.

The central question of this analysis is: **How do the two poets of
resistance, Darwish and Baraka, compare in their emotional expression,
lexical diversity, and structural patterns?**

A bilingual corpus poses methodological challenges, since Arabic and
English words cannot be directly compared. We address this by selecting
techniques that either produce language-independent numeric outputs,
like NRC emotion scores, TTR, and sparsity. Additionally, we run some
techniques, like word frequency, bigram networks, TF-IDF, and pairwise
correlation, separately for each poet and compare them side-by-side.

Our corpus contains **14 poems per poet** (28 total), drawn from
different stages of their careers to reflect changes in style,
vocabulary, and thematic focus over time. The poems were chosen based on
availability, representativeness, and variation in length and style,
rather than through random sampling, meaning the corpus should be
understood as a curated sample rather than a complete representation of
either poet’s work. This approach allows for meaningful comparative
analysis while keeping the dataset manageable for computational
analysis. Darwish’s poems were analyzed in the original Arabic with
orthographic normalization applied, while Baraka’s poems were collected
from publicly available sources and formatted consistently for
line-level analysis.

We apply eight text mining techniques: NRC sentiment analysis,
poem-emotion heatmaps, TF-IDF, lexical diversity (TTR, MATTR), sparsity,
word frequency, bigram networks, and pairwise word correlation.

## 1.1 Background

Mahmoud Darwish is one of the most eminent Arab poets throughout
history. Darwish grew up in a palestinian village, which was later
occupied by Israel. Consequently, he and his family had to flee to
Lebanon, before they could return back to their land. After his return,
Darwish was imprisoned for travelling between villages and reciting
poetry. He therefore lived in exile for large parts of his life. One of
his most popular poems "Identity card" was even made in to a protest
song used for resisting the Israeli military. Darwish was also
politically active, and served as an editor for the Palestinan
Liberation Organisation. This painted the picture of him as a resistance
poet, by strongly fighting and condemning Israeli occupation through
poetry. His background and history naturally shaped his way of writing
and his poems are therefore deeply rooted in the experience of exile,
displacement, and the ongoing struggle for Palestinian identity and
self-determination.

Amiri Baraka was similarly a politically engaged poet and activist. In
contrast to Mahmoud Darwish, Baraka emerged as a central figure in
American literature. He was closely associated with the Black Power
Movement through his role in the Black Arts Movement, which functioned
as its cultural and artistic branch. This movement promoted Black
self-determination, cultural pride, and resistance to systemic racism.
Like Palestinian resistance movements, the Black Power Movement sought
civil rights and autonomy in the face of oppression. Baraka used
literature, theater, and other art forms as tools of political
expression, opposing structures of racial injustice and advocating for
Black liberation.

While both Darwish and Baraka used poetry as a form of resistance, their
approaches differ significantly in tone and style. Darwish's poems use
symbolism and memory to express themes of identity and resistance in a
reflective and lyrical way. Conversely, Baraka's poetry is more direct
and often seen as controversial. By using provocative language and
rhetoric, he directly challenged the system of racial injustice and had
clear demands for change. Moreover, Darwish writes in Arabic, while
Baraka writes in English. The structural and semantic differences
between these two languages can shape how meaning, emotion, and
resistance are expressed. These contrasts highlights the different ways
of expressing resistance which provides an interesting baseline for for
comparing how resistance is linguistically constructed across the two
corpora.

# 2. Setup and Data

## 2.1 Packages

```{r packages}
library(tidyverse)
library(tidytext)
library(textdata)
library(fmsb)
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(wordcloud)
library(reshape2)
library(igraph)
library(ggraph)
library(widyr)
library(RColorBrewer)
library(scales)
library(knitr)
library(ggwordcloud)
library(patchwork)
library(topicmodels)
```

## 2.2 Load Poem Data

```{r load-data}
source("darwish_dataset.R")
source("Baraka_dataset.R")

# Harmonize: add poem_id to Baraka, add English poet name to both
baraka_poems <- baraka_poems %>%
  mutate(poem_id = row_number(),
         poet_en = "Amiri Baraka")

darwish_poems <- darwish_poems %>%
  mutate(poet_en = "Mahmoud Darwish")

# Combined tibble for cross-poet numeric comparisons
all_poems <- bind_rows(
  darwish_poems %>% select(poet_en, poem_id, poem, text),
  baraka_poems  %>% select(poet_en, poem_id, poem, text)
)

cat("Darwish poems:", nrow(darwish_poems), "\n")
cat("Baraka poems:", nrow(baraka_poems), "\n")
```

## 2.3 Stopword Preparation

Arabic stopwords must be normalized with the same function used on the
poems (diacritics removal, alif/ya/ta-marbuta normalization). We also
add common Arabic function words missing from the default list and
filter out onomatopoeia fragments from Baraka's poems.

```{r stopwords}
# Arabic normalization (same as clean_poem in darwish_dataset.R)
normalize_ar <- function(text) {
  text <- stringr::str_remove_all(text, "[\\u064B-\\u065F\\u0670]")
  text <- stringr::str_replace_all(text, "ـ", "")
  text <- stringr::str_replace_all(text, "[إأآٱ]", "ا")
  text <- stringr::str_replace_all(text, "ى", "ي")
  text <- stringr::str_replace_all(text, "ة", "ه")
  text <- stringr::str_replace_all(text, "ؤ", "و")
  text <- stringr::str_replace_all(text, "ئ", "ي")
  text
}

# Normalize the default Arabic stopwords
ar_stops_base <- normalize_ar(stopwords::stopwords("ar", source = "misc"))

# Extended custom Arabic stopwords (common function words missing from default)
ar_stops_custom <- normalize_ar(c(
  "على", "أنا", "يا", "أن", "ما", "لا", "إلى", "هذا", "هذه", "ذلك",
  "التي", "الذي", "كان", "كانت", "عن", "بين", "حين", "بعد", "قبل",
  "فوق", "تحت", "حتى", "إذا", "لكن", "ولكن", "أو", "ثم", "كي",
  "لكي", "منه", "منها", "فيه", "فيها", "عليه", "عليها", "إنه", "إنها",
  "أنت", "أنتِ", "نحن", "هم", "لي", "لنا", "لهم", "بها", "به",
  "كل", "بعض", "كيف", "أين", "متى", "لماذا", "هل", "قد", "سوف",
  "كأن", "مثل", "أيها", "وراء", "أمام", "إلا", "أما",
  "ولم", "لكنني", "ولكنني", "كنت", "كنا", "كان", "ليس",
  "الذين", "اللذين", "واو", "فاء"
))

ar_stops <- unique(c(ar_stops_base, ar_stops_custom))
cat("Total Arabic stopwords (normalized):", length(ar_stops), "\n")

# Custom Baraka stopwords (onomatopoeia fragments)
baraka_custom_stops <- c("bu", "de", "daaaa", "bap", "boo", "bup",
                         "pu", "daaaa", "rrrrrrrrrrrrrrr", "rrrrrrrr",
                         "tuhtuhtuhtuhtuhtuhtuhtuhtuhtuh", "aggh",
                         "whoooo", "whooooooooooooooooooooo")

# Unified English stopwords (tidytext + stopwords package + custom noise)
en_stops <- unique(c(stop_words$word, stopwords::stopwords("en"), baraka_custom_stops))
```

## 2.4 Load NRC Multilingual Emotion Lexicon

The NRC Emotion Lexicon (Mohammad & Turney, 2013) maps words in 100+
languages to 8 emotions (anger, anticipation, disgust, fear, joy,
sadness, surprise, trust) plus positive/negative sentiment. Because both
Arabic and English words are mapped to the same emotion categories, this
serves as the bridge for cross-language comparison.

```{r nrc-lexicon}
nrc_raw <- read_tsv(
  "NRC-Emotion-Lexicon/NRC-Emotion-Lexicon-ForVariousLanguages.txt",
  show_col_types = FALSE
)

emotion_cols <- c("anger", "anticipation", "disgust", "fear",
                  "joy", "negative", "positive", "sadness",
                  "surprise", "trust")

# English NRC lexicon
nrc_english <- nrc_raw %>%
  select(`English Word`, all_of(emotion_cols)) %>%
  pivot_longer(all_of(emotion_cols),
               names_to = "sentiment", values_to = "value") %>%
  filter(value == 1) %>%
  select(word = `English Word`, sentiment) %>%
  distinct()

# Arabic NRC lexicon
nrc_arabic <- nrc_raw %>%
  select(Arabic, all_of(emotion_cols)) %>%
  pivot_longer(all_of(emotion_cols),
               names_to = "sentiment", values_to = "value") %>%
  filter(value == 1) %>%
  select(word = Arabic, sentiment) %>%
  distinct()

cat("English NRC entries:", nrow(nrc_english), "\n")
cat("Arabic NRC entries:", nrow(nrc_arabic), "\n")
```

## 2.5 Corpus Overview

```{r corpus-overview}
corpus_stats <- all_poems %>%
  mutate(
    word_count = str_count(text, "\\S+"),
    line_count = str_count(text, "\n") + 1
  ) %>%
  group_by(poet_en) %>%
  summarise(
    poems = n(),
    total_words = sum(word_count),
    mean_words = round(mean(word_count)),
    median_words = round(median(word_count)),
    total_lines = sum(line_count),
    .groups = "drop"
  )

kable(corpus_stats,
      col.names = c("Poet", "Poems", "Total Words", "Mean Words",
                     "Median Words", "Total Lines"),
      caption = "Corpus summary statistics")
```

# 3. Analysis

## 3.1 NRC Sentiment and Emotion Analysis

This technique assigns emotion labels to every word in each poem using
the NRC Emotion Lexicon. Because the lexicon maps both Arabic and
English to the same emotions, we can directly compare emotional profiles
across languages. We tokenize each poet's poems separately, join to the
appropriate language lexicon, count emotion matches, and normalize by
total matched words per poem to account for length differences.

```{r sentiment-tokenize}
# Tokenize and join to NRC -- Darwish (Arabic)
darwish_tokens_nrc <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc_arabic, by = "word", relationship = "many-to-many")

# Tokenize and join to NRC -- Baraka (English)
baraka_tokens_nrc <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc_english, by = "word", relationship = "many-to-many")

# Report match rates
darwish_total <- darwish_poems %>% unnest_tokens(word, text) %>% nrow()
baraka_total  <- baraka_poems %>% unnest_tokens(word, text) %>% nrow()

cat("Darwish: ", nrow(darwish_tokens_nrc), "emotion matches from",
    darwish_total, "total tokens (",
    round(100 * n_distinct(darwish_tokens_nrc$word) /
            n_distinct(darwish_poems %>% unnest_tokens(word, text) %>% pull(word)), 1),
    "% unique words matched)\n")
cat("Baraka: ", nrow(baraka_tokens_nrc), "emotion matches from",
    baraka_total, "total tokens (",
    round(100 * n_distinct(baraka_tokens_nrc$word) /
            n_distinct(baraka_poems %>% unnest_tokens(word, text) %>% pull(word)), 1),
    "% unique words matched)\n")
```

```{r sentiment-aggregate}
# Aggregate emotion counts per poem
darwish_emotions <- darwish_tokens_nrc %>%
  count(poem, sentiment) %>%
  group_by(poem) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(poet = "Mahmoud Darwish")

baraka_emotions <- baraka_tokens_nrc %>%
  count(poem, sentiment) %>%
  group_by(poem) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(poet = "Amiri Baraka")

all_emotions <- bind_rows(darwish_emotions, baraka_emotions)
```

```{r sentiment-barplot, fig.height=7}
# Average emotion proportion by poet
emotion_summary <- all_emotions %>%
  group_by(poet, sentiment) %>%
  summarise(avg_prop = mean(proportion), .groups = "drop")

ggplot(emotion_summary,
       aes(x = reorder(sentiment, avg_prop), y = avg_prop, fill = poet)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Emotional Profile: Darwish vs. Baraka",
       subtitle = "Average proportion across all poems per poet",
       x = "Emotion / Sentiment", y = "Average Proportion",
       fill = "Poet") +
  theme_minimal(base_size = 13)
```

```{r sentiment-facet, fig.height=8}
# Only the 8 core emotions (exclude positive/negative)
core_emotions <- c("anger", "anticipation", "disgust", "fear",
                   "joy", "sadness", "surprise", "trust")

all_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  group_by(poet, sentiment) %>%
  summarise(avg_prop = mean(proportion),
            sd_prop  = sd(proportion), .groups = "drop") %>%
  ggplot(aes(x = sentiment, y = avg_prop, fill = poet)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = pmax(avg_prop - sd_prop, 0),
                    ymax = avg_prop + sd_prop),
                position = position_dodge(width = 0.8), width = 0.25) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Core Emotion Comparison (8 Emotions)",
       subtitle = "Mean +/- SD across poems",
       x = NULL, y = "Proportion", fill = "Poet") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Both poets share sadness and fear as their two dominant emotions, but
the distributions differ. Darwish's profile is led by sadness (13.6%)
and fear (13.1%), with surprise (11.9%) ranking third. This is a pattern
consistent with exile poetry, where loss and uncertainty dominate.
Baraka's top emotion is fear (12.6%), followed by sadness (10.3%) and
anger (10.2%); his anger score is notably close to his sadness,
reflecting a more confrontational register. Darwish's joy score (4.6%)
is roughly half of Baraka's (8.7%), suggesting a darker emotional
baseline in the Arabic corpus. This may seem counterintuitive,
considering Baraka's confrontational way of writing, but may be a
reflection of themes like empowerment and collective identity within the
Black Arts Movement.

Looking at the second plot, it visualizes how emotions like sadness and
surprise, has larger standard deviations across Darwish's poems. This
suggest greater variability of emotions across the corpus, indicating
that these emotions are unevenly distributed and may reflect particular
moments of heightened intensity rather than a consistent emotional
pattern.

### 3.1.2 Poem-Emotion Heatmap

Having analyzed the core vocabulary, themes, and rhetorics, we now move
over to an analysis of emotions in order to detect emotional patterns
and how they vary across individual poems and within each poet's work.

The heatmap shows the emotional proportions for each individual poem,
revealing which poems contribute most strongly to the overall emotional
tone and which ones deviate from the dominant emotional pattern. This
helps identify emotionally distinctive poems and highlights the degree
of emotional variation within each poet’s work.

```{r heatmap-darwish, fig.height=7, fig.width=10}
# Darwish heatmap
darwish_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  mutate(poem = str_trunc(poem, 30)) %>%
  ggplot(aes(x = sentiment, y = reorder(poem, proportion), fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(proportion, 2)), size = 3, color = "grey20") +
  scale_fill_gradient(low = "white", high = "#B22222",
                      labels = percent_format()) +
  labs(title = "Darwish: Emotion Heatmap by Poem",
       subtitle = "Proportion of emotion words in each poem",
       x = NULL, y = NULL, fill = "Proportion") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.grid = element_blank())
```

```{r heatmap-baraka, fig.height=7, fig.width=10}
# Baraka heatmap
baraka_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  mutate(poem = str_trunc(poem, 30)) %>%
  ggplot(aes(x = sentiment, y = reorder(poem, proportion), fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(proportion, 2)), size = 3, color = "grey20") +
  scale_fill_gradient(low = "white", high = "#1565C0",
                      labels = percent_format()) +
  labs(title = "Baraka: Emotion Heatmap by Poem",
       subtitle = "Proportion of emotion words in each poem",
       x = NULL, y = NULL, fill = "Proportion") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.grid = element_blank())
```

The heatmaps show that the corpus-level averages mask wide variation
within each poet. In Darwish's heatmap, "I Am Joseph, O Father" (أنا
يوسف يا أبي) is dominated by surprise (0.50), while "Rita and the Rifle"
(ريتا والبندقية) peaks on sadness (0.33) and anger (0.17). In Baraka's
heatmap, "Black Art" stands out for its high fear (0.16) and anger
(0.13), while "Preface to a Twenty Volume Suicide Note" is highest on
trust (0.21) -- an unexpected result for a poem with "suicide" in the
title. This poem-level granularity matters: it shows that labeling
either poet with a single emotional tag misrepresents the range of their
work.

## 3.2 Lexical Diversity: TTR and MATTR

Type-Token Ratio (TTR) measures the proportion of unique words (types)
relative to total words (tokens) in a text. A higher TTR indicates more
diverse vocabulary, while a lower TTR suggests more repetition. Because
TTR is a pure ratio computed *within* each text, it is
language-independent. This means that a Darwish poem with TTR = 0.80 and
a Baraka poem with TTR = 0.65 are directly comparable.

However, such comparisons should be interpreted with caution, as Arabic
and English differ structurally in ways that affect vocabulary diversity
scores. Arabic morphology operates at two levels. At the derivational
level, a single root generates semantically distinct words. The root
*w-ṭ-n*, central to Darwish's poetry, yields *waṭan* (homeland),
*muwāṭin* (citizen), and *waṭanī* (national), which are three genuine
vocabulary choices. At the inflectional level, the same word takes
multiple grammatical forms without changing meaning. The words *aḥibbu*
(I love), *tuḥibbu* (you love), and *yuḥibbu* (he loves) are all surface
forms of the same verb, with the pronoun collapsed into the word itself
rather than expressed separately as in English.

As a result, raw TTR counts each surface form as a separate type, which
may inflate lexical diversity in Arabic due to inflectional variation
rather than genuine lexical richness. Ideally, this issue would be
addressed through lemmatization, which reduces inflected forms to a
common base form while preserving derivational differences. However,
Arabic lemmatization was beyond the scope of this project, and the
analysis was conducted using Quanteda without morphological
normalization. This should therefore be considered a limitation of the
study, as lexical diversity scores for Arabic texts may be artificially
inflated compared to English texts.

Simple TTR is sensitive to text length, because longer texts tend to
have lower TTR. So, we also compute the Moving-Average TTR (MATTR),
which uses a sliding window to provide a more robust estimate across
texts of varying length.

```{r ttr-compute}
# Build quanteda corpora with appropriate stopwords
darwish_corpus <- corpus(darwish_poems, text_field = "text",
                         docid_field = "poem")
baraka_corpus  <- corpus(baraka_poems, text_field = "text",
                         docid_field = "poem")

# Tokenize and remove stopwords (normalized for Arabic)
darwish_toks <- tokens(darwish_corpus, remove_punct = TRUE,
                       remove_numbers = TRUE) %>%
  tokens_remove(ar_stops) %>%
  tokens_select(min_nchar = 2)

baraka_toks <- tokens(baraka_corpus, remove_punct = TRUE,
                      remove_numbers = TRUE) %>%
  tokens_remove(en_stops) %>%
  tokens_select(min_nchar = 3)

# DFM
darwish_dfm <- dfm(darwish_toks)
baraka_dfm  <- dfm(baraka_toks)

# Lexical diversity — TTR from DFM, MATTR from tokens
darwish_ttr   <- textstat_lexdiv(darwish_dfm, measure = "TTR")
darwish_mattr <- textstat_lexdiv(darwish_toks, measure = "MATTR")
darwish_div   <- left_join(darwish_ttr, darwish_mattr, by = "document") %>%
  mutate(poet = "Mahmoud Darwish")

baraka_ttr   <- textstat_lexdiv(baraka_dfm, measure = "TTR")
baraka_mattr <- textstat_lexdiv(baraka_toks, measure = "MATTR")
baraka_div   <- left_join(baraka_ttr, baraka_mattr, by = "document") %>%
  mutate(poet = "Amiri Baraka")

all_div <- bind_rows(darwish_div, baraka_div)
```

```{r ttr-boxplot, fig.height=5}
# TTR boxplot
ggplot(all_div, aes(x = poet, y = TTR, fill = poet)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2.5) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Lexical Diversity: Type-Token Ratio (TTR)",
       subtitle = "Each point is one poem",
       y = "TTR (unique words / total words)", x = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

```{r mattr-boxplot, fig.height=5}
# MATTR boxplot
ggplot(all_div, aes(x = poet, y = MATTR, fill = poet)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2.5) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Lexical Diversity: Moving-Average TTR (MATTR)",
       subtitle = "More robust to poem length variation (window = 100 tokens)",
       y = "MATTR", x = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

```{r scores-table}

# Per-poem lexical diversity table
poem_div_table <- all_div %>%
  select(poet, document, TTR, MATTR) %>%
  mutate(
    TTR   = round(TTR, 3),
    MATTR = round(MATTR, 3)
  ) %>%
  arrange(poet, document)

kable(poem_div_table,
      col.names = c("Poet", "Poem", "TTR", "MATTR"),
      caption = "Lexical diversity (TTR and MATTR) by poem",
      align = c("l", "l", "r", "r"))
```

```{r ttr-table}
# Summary statistics
div_summary <- all_div %>%
  group_by(poet) %>%
  summarise(
    mean_TTR   = round(mean(TTR, na.rm = TRUE), 3),
    median_TTR = round(median(TTR, na.rm = TRUE), 3),
    sd_TTR     = round(sd(TTR, na.rm = TRUE), 3),
    mean_MATTR   = round(mean(MATTR, na.rm = TRUE), 3),
    median_MATTR = round(median(MATTR, na.rm = TRUE), 3),
    sd_MATTR     = round(sd(MATTR, na.rm = TRUE), 3),
    .groups = "drop"
  )

kable(div_summary,
      col.names = c("Poet", "Mean TTR", "Median TTR", "SD TTR",
                     "Mean MATTR", "Median MATTR", "SD MATTR"),
      caption = "Lexical diversity summary by poet")
```

The lexical diversity results show generally high TTR and MATTR values
for both poets, indicating that most poems contain a relatively high
proportion of unique words and therefore a diverse vocabulary.

Darwish's mean TTR (0.805) slightly exceeds Baraka's (0.796). Looking at
the MATTR, Baraka's mean MATTR (0.879) is higher than Darwish's (0.854),
suggesting comparable diversity once text length is controlled for. The
outliers are more telling than the averages. Baraka's "SOS" has a TTR of
just 0.27, driven by repetitive incantation from phrases like "Calling
all black people" and "Come in". Darwish's lowest-TTR poem is "أرى شبحي
قادماً من بعيد" / "I See My Ghost Coming from Afar" (0.63), where the
deliberate repetition of the key phrase "I gaze upon" (translated), at
the beginning of each successive clause is a rhetorical strategy, and
known as a common style in Arabic poetry.

Overall, there is no clear overall difference between the two poets
based on TTR and MATTR alone. Variation within each poet’s poems is
often larger than the difference between the poets. This suggests that
lexical diversity is influenced more by individual poem style,
repetition, and structure than by the poet or language alone.

However, as discussed earlier, Arabic morphological structure may
inflate lexical diversity scores. Therefore, Darwish’s lexical diversity
scores may appear slightly higher than they would under lemmatization,
and comparisons between Arabic and English poems should be interpreted
cautiously. For this reason, the results are best interpreted as
comparative within the corpus rather than as absolute measures of
lexical richness across the two languages.

## 3.3 Sparsity and DFM Analysis

While, Lexical Diversity measured the vocabulary variation within
individual poems, Document–Feature Matrix (DFM) analysis allows us to
examine vocabulary distribution across the corpus as a whole. With this
approach, we look at how vocabulary is distributed across all poems by
each poet, particularly how much vocabulary is shared between poems
versus how much is unique to individual poems.

The most important measure here is sparsity which measures the
proportion of zeros in a DFM. High sparsity means most words appear in
very few poems, indicating a vocabulary that is spread thinly across the
corpus. Lower sparsity means more shared vocabulary across poems. This
metric is computed within each poet's sub-corpus, making it comparable
across languages.

```{r sparsity-compute}
# Overall sparsity per poet
darwish_sparsity <- round(sparsity(darwish_dfm) * 100, 1)
baraka_sparsity  <- round(sparsity(baraka_dfm) * 100, 1)

# DFM dimensions
dfm_stats <- tibble(
  Poet = c("Mahmoud Darwish", "Amiri Baraka"),
  Documents = c(ndoc(darwish_dfm), ndoc(baraka_dfm)),
  Features  = c(nfeat(darwish_dfm), nfeat(baraka_dfm)),
  `Features per Doc` = round(c(nfeat(darwish_dfm) / ndoc(darwish_dfm),
                                nfeat(baraka_dfm) / ndoc(baraka_dfm))),
  `Sparsity (%)` = c(darwish_sparsity, baraka_sparsity)
)

kable(dfm_stats,
      col.names = c("Poet", "Documents", "Features", "Features per Doc", "Sparsity (%)"),
      caption = "Document-Feature Matrix comparison")
```

```{r sparsity-trim, fig.height=5}
# Sparsity at different trim levels
trim_levels <- seq(0, 0.5, by = 0.05)

sparsity_by_trim <- map_dfr(trim_levels, function(min_df) {
  d_trim <- dfm_trim(darwish_dfm, min_docfreq = max(min_df, 0.01),
                     docfreq_type = "prop")
  b_trim <- dfm_trim(baraka_dfm, min_docfreq = max(min_df, 0.01),
                     docfreq_type = "prop")
  tibble(
    min_docfreq = min_df,
    Darwish_sparsity = ifelse(nfeat(d_trim) > 0, sparsity(d_trim), NA_real_),
    Baraka_sparsity  = ifelse(nfeat(b_trim) > 0, sparsity(b_trim), NA_real_),
    Darwish_features = nfeat(d_trim),
    Baraka_features  = nfeat(b_trim)
  )
})

sparsity_by_trim %>%
  select(min_docfreq, Darwish_sparsity, Baraka_sparsity) %>%
  pivot_longer(-min_docfreq, names_to = "poet", values_to = "sparsity") %>%
  mutate(poet = str_remove(poet, "_sparsity")) %>%
  ggplot(aes(x = min_docfreq, y = sparsity, color = poet)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(labels = percent_format()) +
  labs(title = "Sparsity vs. Minimum Document Frequency Threshold",
       subtitle = "How sparsity changes as rare terms are progressively removed",
       x = "Minimum Document Frequency (proportion of poems)",
       y = "Sparsity", color = "Poet") +
  theme_minimal(base_size = 13)
```

```{r topfeatures}
# Top features per poet as formatted tables
darwish_top15 <- topfeatures(darwish_dfm, n = 15)
baraka_top15  <- topfeatures(baraka_dfm, n = 15)

kable(
  tibble(Word = names(darwish_top15), Count = unname(darwish_top15)),
  caption = "Top 15 features: Darwish (after stopword removal)"
)
kable(
  tibble(Word = names(baraka_top15), Count = unname(baraka_top15)),
  caption = "Top 15 features: Baraka (after stopword removal)"
)
```

The plot shows how much of a poet's vocabulary is unique to a few poems
versus how much is shared across their entire work. On the far left of
the graph, the high starting points, above 90%, reveal that both Darwish
and Baraka have high sparsity, meaning that the vast majority of words
they use appear in only one or two poems rather than being repeated
throughout the collection. This confirms the finding from Lexical
Diversity analysis, that both authors use a rich and diverse language.
Darwish's DFM has a slightly higher sparsity (92.1%) than Baraka
(91.4%), meaning that a given word in Darwish's corpus is less likely to
appear across multiple poems.

The difference between Sparsity and TTR, is that sparsity looks at the
corpus as a whole, while TTR captures the essence of individual poems.
For instance it includeds deliberate rhetorical repetitions like
Baraka’s incantations or Darwish’s anaphora. The drop in the lines as
the filter moves to the right along the x-axis, shows that once these
unique, poem-specific words are removed, only a very small core of
shared thematic words remains. These words are visible in the top
features. For Baraka, words like "black," "people," and "killed" act as
the connective tissue across his work, while Darwish relies on elemental
imagery such as "sea" (البحر), "land" (الارض), and "sky" (السماء) to
bridge his poems.

The fact that the blue line for Darwish sits higher than the red line
for Baraka suggests that Darwish's vocabulary is even more spread out.
However, this is largely again because of the Arabic language
morphology. In short, the plot illustrates that while these poets have a
few central themes they return to, they primarily use a wide and diverse
range of unique language to express their resistance.

## 3.4 Word Frequency Analysis

As the sparsity of each corpus suggests, the poems have a high
proprotion of unique words, while still being connected through a small
set of recurring themes. Word frequency analysis counts the most common
words after removing stopwords, revealing the core thematic vocabulary
of each poet. Since Arabic and English words cannot be compared
directly, we display each poet's top words separately in a side-by-side
visualization.

```{r word-freq}
# Darwish word frequencies (using normalized Arabic stopwords)
darwish_words <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% ar_stops,
         str_length(word) > 1) %>%
  count(word, sort = TRUE)

# Baraka word frequencies (using unified English stopwords)
baraka_words <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% en_stops,
         str_length(word) > 2) %>%
  count(word, sort = TRUE)
```

```{r freq-barplot, fig.height=7}
top_d <- darwish_words %>% slice_max(n, n = 15) %>%
  mutate(poet = "Mahmoud Darwish")
top_b <- baraka_words %>% slice_max(n, n = 15) %>%
  mutate(poet = "Amiri Baraka")

bind_rows(top_d, top_b) %>%
  ggplot(aes(x = reorder_within(word, n, poet), y = n, fill = poet)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poet, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Top 15 Most Frequent Words by Poet",
       subtitle = "After stopword removal",
       x = NULL, y = "Count") +
  theme_minimal(base_size = 13)
```

```{r wordclouds, fig.height=10, fig.width=16}
wc_data <- bind_rows(
  darwish_words %>% slice_max(n, n = 80) %>% mutate(poet = "Mahmoud Darwish"),
  baraka_words  %>% slice_max(n, n = 80) %>% mutate(poet = "Amiri Baraka")
)

ggplot(wc_data, aes(label = word, size = n, color = n)) +
  geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle") +
  scale_size_area(max_size = 24) +
  scale_color_gradient(low = "darkgreen", high = "darkred") +
  facet_wrap(~poet, scales = "free") +
  theme_minimal(base_size = 14) +
  labs(title = "Word Clouds: Darwish vs. Baraka",
       subtitle = "Top 80 words after stopword removal")
```

The core vocabulary across each poem is what can tell us something about
the themes in each poets writing.

Darwish's most frequent content words after stopword removal are "look
out" (اطل), "and I" (وانا), "my father" (ابي), "the sea" (البحر), and
"and you" (وانت) . This is vocabulary rooted in the self, landscape,
family, and address. It reflects blabla

Baraka's top words are "black," "people," "killed," "poems," "world" ,
reflecting a vocabulary oriented toward collective identity and
violence.

The contrast is stark: Darwish's lexicon is intimate and geographic,
while Baraka's is communal and confrontational.

## 3.5 Bigram Network Graph

Bigram analysis extracts consecutive word pairs and visualizes them as
network graphs, where nodes represent words and edges connect words that
appear together. As with word frequency, bigrams are language-specific
and must be analyzed separately per poet.

This method is used because it provides deeper insight than single words
alone, revealing phrase structure and rhetorical patterns. While a
single-word frequency count might show that a poet uses the word "black"
or "land" often, bigrams reveal the specific context. They can show
whether "black" is paired with "power," "people," or "art," or whether
"land" is paired with "lost," "my," or "holy." By visualizing these
connections as a network, we can identify the "thematic clusters" that
define each poet’s unique style of resistance, allowing us to see not
just their core vocabulary, but the specific associations and meanings
they build through word pairings.

```{r bigrams}
# Darwish bigrams (normalized Arabic stopwords)
darwish_bigrams <- darwish_poems %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% ar_stops, !word2 %in% ar_stops,
         str_length(word1) > 1, str_length(word2) > 1) %>%
  count(word1, word2, sort = TRUE)

# Baraka bigrams (unified English stopwords)
baraka_bigrams <- baraka_poems %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% en_stops, !word2 %in% en_stops,
         str_length(word1) > 2, str_length(word2) > 2) %>%
  count(word1, word2, sort = TRUE)

cat("Darwish unique bigrams:", nrow(darwish_bigrams), "\n")
cat("Baraka unique bigrams:", nrow(baraka_bigrams), "\n")
```

```{r bigram-table}
kable(head(darwish_bigrams, 10),
      col.names = c("Word 1", "Word 2", "Count"),
      caption = "Top 10 Darwish bigrams (after stopword removal)")
kable(head(baraka_bigrams, 10),
      col.names = c("Word 1", "Word 2", "Count"),
      caption = "Top 10 Baraka bigrams (after stopword removal)")
```

```{r bigram-network, fig.height=8, fig.width=14}
# Network plotting function
plot_bigram_network <- function(bigram_counts, title_text,
                                 n_edges = 25, min_n = 2) {
  graph_data <- bigram_counts %>%
    filter(n >= min_n) %>%
    slice_max(n, n = n_edges)

  if (nrow(graph_data) == 0) {
    # If no bigrams meet threshold, lower it
    graph_data <- bigram_counts %>% slice_max(n, n = n_edges)
  }

  graph <- graph_from_data_frame(graph_data)

  ggraph(graph, layout = "fr") +
    geom_edge_link(
      aes(edge_alpha = n, edge_width = n),
      show.legend = FALSE,
      arrow = arrow(type = "closed", length = unit(0.12, "inches")),
      end_cap = circle(3, "mm")
    ) +
    geom_node_point(color = "steelblue", size = 5, alpha = 0.8) +
    geom_node_text(aes(label = name), vjust = 1.2, hjust = 0.5,
                   size = 3.5, check_overlap = TRUE) +
    theme_void(base_size = 13) +
    labs(title = title_text)
}

# Side-by-side networks
set.seed(42)
p_darwish <- plot_bigram_network(darwish_bigrams,
                                  "Darwish: Bigram Network",
                                  n_edges = 30, min_n = 2)
p_baraka  <- plot_bigram_network(baraka_bigrams,
                                  "Baraka: Bigram Network",
                                  n_edges = 30, min_n = 2)

p_darwish + p_baraka +
  plot_annotation(
    title = "Bigram Network Graphs: Darwish vs. Baraka",
    subtitle = "Nodes = words, edges = consecutive word pairs (after stopword removal)",
    theme = theme(plot.title = element_text(size = 16, face = "bold"))
  )
```

Baraka's top bigram is "black people," which functions as both subject
and rallying cry across multiple poems. His network also shows the
striking pair "survive blackman / blackman survive" -- a chiastic
repetition typical of his incantatory style. Darwish's top bigrams
include "think of others" (فكر بغيرك, also a poem title used as a
refrain) and "passing words" (الكلمات العابره). Where Baraka's network
clusters tightly around racial identity, Darwish's spreads across
landscape and temporal imagery.

## 3.8 TF-IDF: Distinctive Words

Term Frequency-Inverse Document Frequency (TF-IDF) identifies words that
are distinctive to each poet rather than just frequent. A word with high
TF-IDF appears often in one poet's work but rarely, or never, in the
other's. This is computed by treating each poet's full corpus as a
single document. Since Arabic and English never share words, TF-IDF is
applied within each language separately across poems, in order to
analyze which words are distinctive to each poem and poet across the
corpus.

Four poems are chosen from each poet to show the TF-IDF per poem for
those. write more about why

```{r tfidf, fig.height=8, fig.width=12}
# TF-IDF per poem within each poet
darwish_tfidf <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% ar_stops, str_length(word) > 1) %>%
  count(poem, word, sort = TRUE) %>%
  bind_tf_idf(word, poem, n) 

baraka_tfidf <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% en_stops, str_length(word) > 2) %>%
  count(poem, word, sort = TRUE) %>%
  bind_tf_idf(word, poem, n)

# Top TF-IDF words per poet (across all poems)
top_tfidf_d <- darwish_tfidf %>%
  arrange(desc(tf_idf)) %>%
  distinct(word, .keep_all = TRUE) %>%
  slice_max(tf_idf, n = 15) %>%
  mutate(poet = "Mahmoud Darwish")

top_tfidf_b <- baraka_tfidf %>%
  arrange(desc(tf_idf)) %>%
  distinct(word, .keep_all = TRUE) %>%
  slice_max(tf_idf, n = 15) %>%
  mutate(poet = "Amiri Baraka")

bind_rows(top_tfidf_d, top_tfidf_b) %>%
  ggplot(aes(x = reorder_within(word, tf_idf, poet),
             y = tf_idf, fill = poet)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poet, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Most Distinctive Words by Poet (TF-IDF)",
       subtitle = "Words that are uniquely important to individual poems",
       x = NULL, y = "TF-IDF Score") +
  theme_minimal(base_size = 13)
```

```{r tfidf-per-poem, fig.height=9, fig.width=13}
# Top TF-IDF words per poem -- Baraka (4 poems)
baraka_tfidf %>%
  filter(poem %in% c("Somebody Blew Up America", "Black Art",
                      "Preface to a Twenty Volume Suicide Note",
                      "An Agony. As Now.")) %>%
  group_by(poem) %>%
  slice_max(tf_idf, n = 8) %>%
  ungroup() %>%
  ggplot(aes(x = reorder_within(word, tf_idf, poem),
             y = tf_idf, fill = poem)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poem, scales = "free", ncol = 2) +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Baraka: Most Distinctive Words per Poem (TF-IDF)",
       subtitle = "Words that define each poem's unique vocabulary",
       x = NULL, y = "TF-IDF") +
  theme_minimal(base_size = 12)
```

```{r td-idf-darwish, fig.height=9, fig.width=13}
# Top TF-IDF words per poem -- Darwish (4 poems)
darwish_tfidf %>%
  filter(poem %in% c("بطاقة هوية", "ريتا والبندقية",
                      "إلى أمي",
                      "أرى شبحي قادماً من بعيد")) %>%
  group_by(poem) %>%
  slice_max(tf_idf, n = 8) %>%
  ungroup() %>%
  ggplot(aes(x = reorder_within(word, tf_idf, poem),
             y = tf_idf, fill = poem)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poem, scales = "free", ncol = 2) +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Darwish: Most Distinctive Words per Poem (TF-IDF)",
       subtitle = "Words that define each poem's unique vocabulary",
       x = NULL, y = "TF-IDF") +
  theme_minimal(base_size = 12)
```

Where frequency analysis shows what is common across a poet's work,
TF-IDF shows what makes each poem unique. The per-poem TF-IDF plots for
Baraka confirm that "Somebody Blew Up America" is dominated by
accusatory vocabulary, while "Preface" is defined by more introspective
terms. These thematic signatures help distinguish poems that share
high-frequency words but differ in subject and tone.

# 4. Discussion

## 4.1 Emotional Registers of Resistance

The NRC emotion analysis is the most direct cross-linguistic comparison
in this study. Both poets share fear and sadness as their dominant
emotions, confirming that resistance poetry -- whether rooted in
Palestinian exile or African-American struggle -- draws heavily on these
registers. The difference lies in the secondary emotions: Darwish leans
toward surprise (11.9%), reflecting the uncertainty and displacement of
exile, while Baraka's anger (10.2%) runs nearly as strong as his
sadness, consistent with his confrontational poetics.

One caveat: not every word matches the NRC lexicon. Arabic's
morphological complexity reduces match rates because prefixes and
suffixes create surface forms absent from the lexicon. We report match
rates transparently so the reader can judge the reliability of these
profiles.

## 4.2 Vocabulary and Structure

Mean TTR values are close (Darwish 0.805, Baraka 0.796), but the
outliers tell a richer story. Baraka's "SOS" has a TTR of just 0.27 --
the lowest in either corpus -- because its heavy onomatopoeia and
incantatory repetition deliberately sacrifice vocabulary breadth for
rhythmic force. Darwish uses anaphora similarly in "Passers in Passing
Words" (TTR 0.63), where the repeated address structure is a rhetorical
choice, not a vocabulary limitation. In both cases, low TTR flags
deliberate poetic strategy.

Sparsity is comparable across the two sub-corpora (Darwish 92.1%, Baraka
91.4%), indicating that both poets draw on largely poem-specific
vocabulary rather than recycling a fixed set of terms. This high
sparsity is expected in poetry, where each poem typically addresses a
distinct theme or occasion.

## 4.3 Thematic Vocabulary and Phrase Patterns

The word frequency and bigram analyses run separately per language but
expose a clear thematic contrast. Darwish's top words (اطل "look out,"
ابي "my father," البحر "the sea") are intimate and landscape-oriented.
Baraka's ("black," "people," "killed") are collective and political. The
bigram "black people" dominates Baraka's network as both subject and
rallying cry, while Darwish's top bigrams ("think of others," "passing
words") are more reflective.

Word frequency and bigrams complement each other: frequency shows what
each poet returns to, while bigrams show how those terms combine into
phrases. Baraka's chiastic pair "survive blackman / blackman survive" is
invisible in a word count but central to his rhythmic strategy.

## 4.4 Cross-Linguistic Patterns

The poem-level heatmaps show that these aggregate profiles mask wide
internal variation. Darwish's "I Am Joseph, O Father" is dominated by
surprise (0.50), while "Rita and the Rifle" peaks on sadness and anger.
Baraka's "Black Art" concentrates on fear and anger, while "Preface to a
Twenty Volume Suicide Note" scores highest on trust (0.21). Neither poet
is a one-note writer, even though both are often filed under
"resistance."

TF-IDF and the per-poem breakdown shift attention from shared vocabulary
to distinctive vocabulary. Words that define "Somebody Blew Up America"
differ sharply from those that define "Preface," even though both poems
share many high-frequency terms. This technique helps separate thematic
identity from mere frequency.

Pairwise correlation networks group words that share thematic space
across poems, regardless of adjacency. Baraka's network clusters racial
identity vocabulary separately from political accusation vocabulary.
Darwish's clusters landscape imagery with body imagery. These groupings
map the latent thematic modules that organize each poet's work.

## 4.5 Limitations

We note several methodological limitations:

1.  **Arabic morphology**: Arabic's rich inflectional system produces
    many surface forms from a single root. Without lemmatization or
    stemming (which would require specialized Arabic NLP tools beyond
    the scope of this course), each inflected form is treated as a
    separate word. This may inflate TTR and sparsity for Darwish
    relative to Baraka, and reduce NRC lexicon match rates.

2.  **Corpus size**: With 14 poems per poet, this is a small corpus by
    text mining standards. Results are exploratory and should not be
    generalized to the full body of either poet's work.

3.  **NRC lexicon coverage**: The multilingual NRC lexicon relies on
    machine translation for non-English entries. Some Arabic emotion
    assignments may be imprecise due to translation artifacts or
    cultural differences in emotional connotation.

4.  **Translation equivalence**: We compare numeric metrics, not
    meanings. Two poems might score identically on "sadness" for
    entirely different reasons, rooted in different cultural contexts of
    loss.

# 5. Conclusion

Darwish and Baraka share sadness and fear as their dominant emotions,
but diverge on secondary registers: Darwish's work carries more surprise
and less joy, while Baraka balances anger with trust and affirmation.
Their lexical strategies also differ -- Darwish's vocabulary is intimate
and geographic, Baraka's is collective and confrontational -- though
both poets show high within-corpus variation, with individual poems
occupying very different emotional and thematic positions.

The ten techniques applied here -- NRC sentiment, heatmaps, TF-IDF,
TTR/MATTR, sparsity, word frequency, bigrams, and pairwise correlation,
covers tokenization, sentiment analysis, lexical diversity,
document-feature matrices, and network analysis. The NRC multilingual
lexicon proved essential for bridging the language divide, enabling
direct comparison of emotional profiles between Arabic and English
poetry.

The poem-level analyses (heatmaps, TF-IDF) are where the most
interesting patterns emerge: they show that neither poet is monolithic,
and that corpus-level averages obscure substantial internal variation.
These quantitative patterns complement traditional literary analysis by
making visible the emotional and structural architecture of resistance
poetry across two very different linguistic traditions.

# References

-   Benoit, K. et al. (2018). quanteda: An R package for the
    quantitative analysis of textual data. *Journal of Open Source
    Software*, 3(30), 774.
-   Covington, M. A. & McFall, J. D. (2010). Cutting the Gordian Knot:
    The Moving-Average Type-Token Ratio (MATTR). *Journal of
    Quantitative Linguistics*, 17(2), 94-100.
-   Mohammad, S. M. & Turney, P. D. (2013). Crowdsourcing a Word-Emotion
    Association Lexicon. *Computational Intelligence*, 29(3), 436-465.
-   Silge, J. & Robinson, D. (2017). *Text Mining with R: A Tidy
    Approach*. O'Reilly.
-   Welbers, K. et al. (2017). Text Analysis in R. *Communication
    Methods and Measures*, 11(4), 245-265.
-   Poetry Foundation. (n.d.). Mahmoud Darwish.
    <https://www.poetryfoundation.org/poets/mahmoud-darwish>
-   Poetry Foundation. (n.d.). Amiri Baraka.
    <https://www.poetryfoundation.org/poets/amiri-baraka>
-   Poem sources: aldiwan.net, poetsgate.com, poetryfoundation.org,
    poemhunter.com, Genius, becauseofthemwecan.com

------------------------------------------------------------------------

*Analysis conducted with R `r getRversion()`, using tidytext, quanteda,
and the NRC Emotion Lexicon.*

```{r session-info}
sessionInfo()
```
