---
title: "The Poetry of Resistance: A Cross-Linguistic Text Mining Analysis"
subtitle: "Mahmoud Darwish (Arabic) vs. Amiri Baraka (English)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: flatly
    code_folding: hide
    css: style.css
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 6, fig.align = "center",
  dev = if (requireNamespace("ragg", quietly = TRUE)) "ragg_png" else "png",
  dpi = 150
)
```

:::::::::::::::::: project-header
:::: team-card
<h4>Team</h4>

<ul>

<li>[Abdullah Tadmuri]{.member-name}</li>

<li>[Maiheliya Maimaitimin]{.member-name}</li>

<li>[Tone Varberg Sabri]{.member-name}</li>

</ul>

::: course-line
Text Mining Final Project · UC3M · 2025–2026
:::
::::

::::::::::::::: stats-grid
::::: stat-card
::: stat-number
28
:::

::: stat-label
Poems Analyzed
:::
:::::

::::: stat-card
::: stat-number
6
:::

::: stat-label
Techniques Used
:::
:::::

::::: stat-card
::: stat-number
2
:::

::: stat-label
Languages
:::
:::::

::::: stat-card
::: stat-number
14
:::

::: stat-label
Poems per Poet
:::
:::::
:::::::::::::::
::::::::::::::::::

# 1. Introduction

This study analyzes two prominent resistance poets from distinct
traditions: **Mahmoud Darwish** (1941--2008), the Palestinian national
poet who wrote in Arabic, and **Amiri Baraka** (1934--2014), the
African-American poet and activist who wrote in English.

Both poets use poetry as a tool for collective resistance, though they
address different, but comparable struggles. Darwish's poetry centers on
Palestinian dispossession, exile, and the longing for homeland, while
Baraka's work confronts anti-Black racism, celebrates Black identity,
and calls for revolutionary change.

The central question of this analysis is: **How do the two poets of
resistance, Darwish and Baraka, compare in their emotional expression,
lexical diversity, and structural patterns?**

Our corpus contains **14 poems per poet** (28 total), drawn from
different stages of their careers to reflect changes in style,
vocabulary, and thematic focus over time. The poems were chosen based on
availability, representativeness, and variation in length and style,
rather than through random sampling, meaning the corpus should be
understood as a curated sample rather than a complete representation of
either poet’s work. This approach allows for meaningful comparative
analysis while keeping the dataset manageable for computational
analysis. Darwish’s poems were analyzed in the original Arabic with
orthographic normalization applied, while Baraka’s poems were collected
from publicly available sources and formatted consistently for
line-level analysis.

We apply six text mining techniques: NRC sentiment analysis, lexical
diversity (TTR), sparsity, word frequency, bigram networks, and TF-IDF.

## 1.1 Background

Mahmoud Darwish is one of the most eminent Arab poets throughout
history. Darwish grew up in a palestinian village, which was later
occupied by Israel. Consequently, he and his family had to flee to
Lebanon, before they could return back to their land. After his return,
Darwish was imprisoned for travelling between villages and reciting
poetry. He therefore lived in exile for large parts of his life. One of
his most popular poems "Identity card" was even made in to a protest
song used for resisting the Israeli military. Darwish was also
politically active, and served as an editor for the Palestinan
Liberation Organisation. This painted the picture of him as a resistance
poet, by strongly fighting and condemning Israeli occupation through
poetry. His poems are deeply rooted in the experience of exile,
displacement, and the ongoing struggle for Palestinian identity and
self-determination.

Amiri Baraka was similarly a politically engaged poet and activist. In
contrast to Mahmoud Darwish, Baraka emerged as a central figure in
American literature. He was closely associated with the Black Power
Movement through his role in the Black Arts Movement, which functioned
as its cultural and artistic branch. This movement promoted Black
self-determination, cultural pride, and resistance to systemic racism.
Baraka used literature, theater, and other art forms as tools of
political expression, opposing structures of racial injustice and
advocating for Black liberation.

## 1.2 Linguistic Challenges

A bilingual corpus poses methodological challenges, since Arabic and
English words cannot be directly compared. We address this by selecting
techniques that produce language-independent numeric outputs, like NRC
emotion scores, MATTR, and sparsity. Additionally, word frequency,
bigram networks, and TF-IDF are run separately for each poet to allow
for a side-by-side comparison.

Arabic and English differ structurally in ways that affect vocabulary
diversity. Arabic morphology operates at two levels. At the derivational
level, a single root generates semantically distinct words; for
instance, the root *w-ṭ-n* (وطن), central to Darwish's poetry, yields
*waṭan* (homeland) and *muwāṭin* (citizen), which are genuine vocabulary
choices. At the inflectional level, the same word takes multiple
grammatical forms without changing meaning. The words *aḥibbu* (أحب, I
love) and *tuḥibbu* (تحب, you love) are surface forms of the same verb,
with the pronoun collapsed into the word itself rather than expressed
separately as in English.

Ideally, this would be addressed through lemmatization to reduce
inflected forms to a common base. However, Arabic lemmatization was
beyond the scope of this project, and the analysis was conducted using
Quanteda without morphological normalization. This is a limitation of
the study, as lexical diversity and sparsity scores for Darwish may be
artificially inflated compared to Baraka.

# 2. Setup and Data

## 2.1 Packages

```{r packages}
library(tidyverse)
library(tidytext)
library(textdata)
library(fmsb)
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(wordcloud)
library(reshape2)
library(igraph)
library(ggraph)
library(widyr)
library(RColorBrewer)
library(scales)
library(knitr)
library(ggwordcloud)
library(patchwork)
library(topicmodels)
```

## 2.2 Load Poem Data

```{r load-data}
source("darwish_dataset.R")
source("Baraka_dataset.R")

# Harmonize: add poem_id to Baraka, add English poet name to both
baraka_poems <- baraka_poems %>%
  mutate(poem_id = row_number(),
         poet_en = "Amiri Baraka")

darwish_poems <- darwish_poems %>%
  mutate(poet_en = "Mahmoud Darwish")

# Combined tibble for cross-poet numeric comparisons
all_poems <- bind_rows(
  darwish_poems %>% select(poet_en, poem_id, poem, text),
  baraka_poems  %>% select(poet_en, poem_id, poem, text)
)

cat("Darwish poems:", nrow(darwish_poems), "\n")
cat("Baraka poems:", nrow(baraka_poems), "\n")
```

## 2.3 Stopword Preparation

Arabic stopwords must be normalized with the same function used on the
poems (diacritics removal, alif/ya/ta-marbuta normalization). We also
add common Arabic function words missing from the default list and
filter out onomatopoeia fragments from Baraka's poems.

```{r stopwords}
# Arabic normalization (same as clean_poem in darwish_dataset.R)
normalize_ar <- function(text) {
  text <- stringr::str_remove_all(text, "[\\u064B-\\u065F\\u0670]")
  text <- stringr::str_replace_all(text, "ـ", "")
  text <- stringr::str_replace_all(text, "[إأآٱ]", "ا")
  text <- stringr::str_replace_all(text, "ى", "ي")
  text <- stringr::str_replace_all(text, "ة", "ه")
  text <- stringr::str_replace_all(text, "ؤ", "و")
  text <- stringr::str_replace_all(text, "ئ", "ي")
  text
}

# Normalize the default Arabic stopwords
ar_stops_base <- normalize_ar(stopwords::stopwords("ar", source = "misc"))

# Extended custom Arabic stopwords (common function words missing from default)
ar_stops_custom <- normalize_ar(c(
  "على", "أنا", "يا", "أن", "ما", "لا", "إلى", "هذا", "هذه", "ذلك",
  "التي", "الذي", "كان", "كانت", "عن", "بين", "حين", "بعد", "قبل",
  "فوق", "تحت", "حتى", "إذا", "لكن", "ولكن", "أو", "ثم", "كي",
  "لكي", "منه", "منها", "فيه", "فيها", "عليه", "عليها", "إنه", "إنها",
  "أنت", "أنتِ", "نحن", "هم", "لي", "لنا", "لهم", "بها", "به",
  "كل", "بعض", "كيف", "أين", "متى", "لماذا", "هل", "قد", "سوف",
  "كأن", "مثل", "أيها", "وراء", "أمام", "إلا", "أما",
  "ولم", "لكنني", "ولكنني", "كنت", "كنا", "كان", "ليس",
  "الذين", "اللذين", "واو", "فاء"
))

ar_stops <- unique(c(ar_stops_base, ar_stops_custom))
cat("Total Arabic stopwords (normalized):", length(ar_stops), "\n")

# Custom Baraka stopwords (onomatopoeia fragments)
baraka_custom_stops <- c("bu", "de", "daaaa", "bap", "boo", "bup",
                         "pu", "daaaa", "rrrrrrrrrrrrrrr", "rrrrrrrr",
                         "tuhtuhtuhtuhtuhtuhtuhtuhtuhtuh", "aggh",
                         "whoooo", "whooooooooooooooooooooo", "heh")

# Unified English stopwords (tidytext + stopwords package + custom noise)
en_stops <- unique(c(stop_words$word, stopwords::stopwords("en"), baraka_custom_stops))
```

## 2.4 Load NRC Multilingual Emotion Lexicon

The NRC Emotion Lexicon (Mohammad & Turney, 2013) maps words in 100+
languages to 8 emotions (anger, anticipation, disgust, fear, joy,
sadness, surprise, trust) plus positive/negative sentiment. Because both
Arabic and English words are mapped to the same emotion categories, this
serves as the bridge for cross-language comparison.

```{r nrc-lexicon}
nrc_raw <- read_tsv(
  "NRC-Emotion-Lexicon/NRC-Emotion-Lexicon-ForVariousLanguages.txt",
  show_col_types = FALSE
)

emotion_cols <- c("anger", "anticipation", "disgust", "fear",
                  "joy", "negative", "positive", "sadness",
                  "surprise", "trust")

# English NRC lexicon
nrc_english <- nrc_raw %>%
  select(`English Word`, all_of(emotion_cols)) %>%
  pivot_longer(all_of(emotion_cols),
               names_to = "sentiment", values_to = "value") %>%
  filter(value == 1) %>%
  select(word = `English Word`, sentiment) %>%
  distinct()

# Arabic NRC lexicon
nrc_arabic <- nrc_raw %>%
  select(Arabic, all_of(emotion_cols)) %>%
  pivot_longer(all_of(emotion_cols),
               names_to = "sentiment", values_to = "value") %>%
  filter(value == 1) %>%
  select(word = Arabic, sentiment) %>%
  distinct()

cat("English NRC entries:", nrow(nrc_english), "\n")
cat("Arabic NRC entries:", nrow(nrc_arabic), "\n")
```

## 2.5 Corpus Overview

```{r corpus-overview}
corpus_stats <- all_poems %>%
  mutate(
    word_count = str_count(text, "\\S+"),
    line_count = str_count(text, "\n") + 1
  ) %>%
  group_by(poet_en) %>%
  summarise(
    poems = n(),
    total_words = sum(word_count),
    mean_words = round(mean(word_count)),
    median_words = round(median(word_count)),
    total_lines = sum(line_count),
    .groups = "drop"
  )

kable(corpus_stats,
      col.names = c("Poet", "Poems", "Total Words", "Mean Words",
                     "Median Words", "Total Lines"),
      caption = "Corpus summary statistics")
```

# 3. Analysis

## 3.1 NRC Sentiment and Emotion Analysis

We begin by using the NRC Lexicon to assign emotion labels to every word
in each poem. We tokenize each poet's poems separately, join to the
appropriate language lexicon, count emotion matches, and normalize by
total matched words per poem to account for length differences.

```{r sentiment-tokenize}
# Tokenize and join to NRC -- Darwish (Arabic)
darwish_tokens_nrc <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc_arabic, by = "word", relationship = "many-to-many")

# Tokenize and join to NRC -- Baraka (English)
baraka_tokens_nrc <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc_english, by = "word", relationship = "many-to-many")

# Report match rates
darwish_total <- darwish_poems %>% unnest_tokens(word, text) %>% nrow()
baraka_total  <- baraka_poems %>% unnest_tokens(word, text) %>% nrow()

cat("Darwish: ", nrow(darwish_tokens_nrc), "emotion matches from",
    darwish_total, "total tokens (",
    round(100 * n_distinct(darwish_tokens_nrc$word) /
            n_distinct(darwish_poems %>% unnest_tokens(word, text) %>% pull(word)), 1),
    "% unique words matched)\n")
cat("Baraka: ", nrow(baraka_tokens_nrc), "emotion matches from",
    baraka_total, "total tokens (",
    round(100 * n_distinct(baraka_tokens_nrc$word) /
            n_distinct(baraka_poems %>% unnest_tokens(word, text) %>% pull(word)), 1),
    "% unique words matched)\n")
```

Darwish produced only 283 emotion matches from 3001 total tokens (4.7%
match rate), while Baraka produced 1125 matches from 4017 tokens
(19.3%). This difference means that Darwish's emotion scores are based
on a much smaller proportion of his vocabulary, likely reflecting
Arabic's morphological, as well as potential imprecision in the NRC's
machine-translated Arabic entries.

```{r sentiment-aggregate}
# Aggregate emotion counts per poem
darwish_emotions <- darwish_tokens_nrc %>%
  count(poem, sentiment) %>%
  group_by(poem) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(poet = "Mahmoud Darwish")

baraka_emotions <- baraka_tokens_nrc %>%
  count(poem, sentiment) %>%
  group_by(poem) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(poet = "Amiri Baraka")

all_emotions <- bind_rows(darwish_emotions, baraka_emotions)
```

```{r sentiment-barplot, fig.height=7}
# Average emotion proportion by poet
emotion_summary <- all_emotions %>%
  group_by(poet, sentiment) %>%
  summarise(avg_prop = mean(proportion), .groups = "drop")

ggplot(emotion_summary,
       aes(x = reorder(sentiment, avg_prop), y = avg_prop, fill = poet)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Emotional Profile: Darwish vs. Baraka",
       subtitle = "Average proportion across all poems per poet",
       x = "Emotion / Sentiment", y = "Average Proportion",
       fill = "Poet") +
  theme_minimal(base_size = 13)
```

```{r sentiment-facet, fig.height=8}
# Only the 8 core emotions (exclude positive/negative)
core_emotions <- c("anger", "anticipation", "disgust", "fear",
                   "joy", "sadness", "surprise", "trust")

all_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  group_by(poet, sentiment) %>%
  summarise(avg_prop = mean(proportion),
            sd_prop  = sd(proportion), .groups = "drop") %>%
  ggplot(aes(x = sentiment, y = avg_prop, fill = poet)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = pmax(avg_prop - sd_prop, 0),
                    ymax = avg_prop + sd_prop),
                position = position_dodge(width = 0.8), width = 0.25) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Core Emotion Comparison (8 Emotions)",
       subtitle = "Mean +/- SD across poems",
       x = NULL, y = "Proportion", fill = "Poet") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Both poets share sadness and fear as dominant emotions, yet their
distributions reveal distinct rhetorical strategies. Darwish’s profile
is led by sadness (13.6%), fear (13.1%), and surprise (11.9%), a pattern
consistent with exile poetry, where loss and uncertainty prevail.
Baraka’s profile shifts toward a more confrontational register, with
anger (10.2%) nearly equaling his sadness.

Thematically, Baraka’s higher joy score (8.7% vs Darwish’s 4.6%) likely
reflects the Black Arts Movement’s emphasis on spiritual celebration and
cultural pride, creating space for joy even within a resistance corpus.
Conversely, Darwish’s structural focus on displacement and identity
leaves less room for joyful registers. While Darwish exhibits higher
overall sentiment polarization (positive/negative), Baraka uses a more
"neutral" political vocabulary that bypasses simple sentiment
classification.

### Poem-Emotion Heatmap

```{r heatmap-darwish, fig.height=7, fig.width=10}
# Darwish heatmap
darwish_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  mutate(poem = str_trunc(poem, 30)) %>%
  ggplot(aes(x = sentiment, y = reorder(poem, proportion), fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(proportion, 2)), size = 3, color = "grey20") +
  scale_fill_gradient(low = "white", high = "#B22222",
                      labels = percent_format()) +
  labs(title = "Darwish: Emotion Heatmap by Poem",
       subtitle = "Proportion of emotion words in each poem",
       x = NULL, y = NULL, fill = "Proportion") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.grid = element_blank())
```

```{r heatmap-baraka, fig.height=7, fig.width=10}
# Baraka heatmap
baraka_emotions %>%
  filter(sentiment %in% core_emotions) %>%
  mutate(poem = str_trunc(poem, 30)) %>%
  ggplot(aes(x = sentiment, y = reorder(poem, proportion), fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(proportion, 2)), size = 3, color = "grey20") +
  scale_fill_gradient(low = "white", high = "#1565C0",
                      labels = percent_format()) +
  labs(title = "Baraka: Emotion Heatmap by Poem",
       subtitle = "Proportion of emotion words in each poem",
       x = NULL, y = NULL, fill = "Proportion") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.grid = element_blank())
```

The most significant contrast between the two poets is structural.
Darwish’s heatmap reveals "emotional concentration"; high intensity is
localized in a few cells. This suggests his poems are emotionally
focused rather than multidimensional. For example, his 13.6% sadness
score is heavily driven by ريتا والبندقية (*Rita and the Rifle* ) at
0.33 and الجدار (*The Wall)* at 0.22. In these works, sadness is a
direct response to the "war-side" of occupation. Similarly, surprise is
almost exclusively tied to أنا يوسف يا أبي (*I Am Joseph, O Father* ) at
0.50. A notable exception is فكر بغيرك (*Think of Others*), which shows
a rare, even distribution (0.17) across four emotions, reflecting a call
for universal empathy.

In contrast, Baraka’s heatmap is "uniformly dense," reflecting a more
consistent emotional engagement across his work. His high anger scores
appears consistently in poems like *Legacy*, *Notes for a Speech*, and
*Black Art*. This confirms that for Baraka, anger is not an isolated
reaction but a permanent rhetorical feature of his resistance. Likewise,
his joy is distributed across introspective works like *Preface to a
Twenty Volume Suicide Note* and *An Agony. As Now*.

Ultimately, these patterns map onto their stylistic differences. Darwish
writes concentrated lyrical poems where a single emotional register
dominates a specific scene, while Baraka’s varied poems distribute
multiple emotions simultaneously to create a broad political critique.

## 3.2 Lexical Diversity:

Type-Token Ratio (TTR) measures the proportion of unique words (types)
relative to total words (tokens) in a text. Considering TTR is sensitive
to text length, we compute and use the Moving-Average TTR (MATTR), which
uses a more robust estimate across texts of varying length.

A higher score indicates more diverse vocabulary, while a lower score
suggests more repetition.

```{r ttr-compute}
# Build quanteda corpora with appropriate stopwords
darwish_corpus <- corpus(darwish_poems, text_field = "text",
                         docid_field = "poem")
baraka_corpus  <- corpus(baraka_poems, text_field = "text",
                         docid_field = "poem")

# Tokenize and remove stopwords (normalized for Arabic)
darwish_toks <- tokens(darwish_corpus, remove_punct = TRUE,
                       remove_numbers = TRUE) %>%
  tokens_remove(ar_stops) %>%
  tokens_select(min_nchar = 2)

baraka_toks <- tokens(baraka_corpus, remove_punct = TRUE,
                      remove_numbers = TRUE) %>%
  tokens_remove(en_stops) %>%
  tokens_select(min_nchar = 3)

# DFM
darwish_dfm <- dfm(darwish_toks)
baraka_dfm  <- dfm(baraka_toks)

# Lexical diversity — TTR from DFM, MATTR from tokens
darwish_ttr   <- textstat_lexdiv(darwish_dfm, measure = "TTR")
darwish_mattr <- textstat_lexdiv(darwish_toks, measure = "MATTR")
darwish_div   <- left_join(darwish_ttr, darwish_mattr, by = "document") %>%
  mutate(poet = "Mahmoud Darwish")

baraka_ttr   <- textstat_lexdiv(baraka_dfm, measure = "TTR")
baraka_mattr <- textstat_lexdiv(baraka_toks, measure = "MATTR")
baraka_div   <- left_join(baraka_ttr, baraka_mattr, by = "document") %>%
  mutate(poet = "Amiri Baraka")

all_div <- bind_rows(darwish_div, baraka_div)
```

```{r mattr-boxplot, fig.height=5}
# MATTR boxplot
ggplot(all_div, aes(x = poet, y = MATTR, fill = poet)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2.5) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Lexical Diversity: Moving-Average TTR (MATTR)",
       subtitle = "More robust to poem length variation (window = 100 tokens)",
       y = "MATTR", x = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

```{r scores-table}

# Per-poem lexical diversity table
poem_div_table <- all_div %>%
  select(poet, document, TTR, MATTR) %>%
  mutate(
    TTR   = round(TTR, 3),
    MATTR = round(MATTR, 3)
  ) %>%
  arrange(poet, document)

kable(poem_div_table,
      col.names = c("Poet", "Poem", "TTR", "MATTR"),
      caption = "Lexical diversity (TTR and MATTR) by poem",
      align = c("l", "l", "r", "r"))
```

```{r ttr-table}
# Summary statistics
div_summary <- all_div %>%
  group_by(poet) %>%
  summarise(
    mean_TTR   = round(mean(TTR, na.rm = TRUE), 3),
    median_TTR = round(median(TTR, na.rm = TRUE), 3),
    sd_TTR     = round(sd(TTR, na.rm = TRUE), 3),
    mean_MATTR   = round(mean(MATTR, na.rm = TRUE), 3),
    median_MATTR = round(median(MATTR, na.rm = TRUE), 3),
    sd_MATTR     = round(sd(MATTR, na.rm = TRUE), 3),
    .groups = "drop"
  )

kable(div_summary,
      col.names = c("Poet", "Mean TTR", "Median TTR", "SD TTR",
                     "Mean MATTR", "Median MATTR", "SD MATTR"),
      caption = "Lexical diversity summary by poet")
```

Both poets uses a rich, non-repetitive vocabulary. While Baraka’s mean
MATTR (0.879) is slightly higher than Darwish’s (0.854), the specific
outliers are more analytically significant than the averages. For
Baraka, *SOS* is a notable outlier (0.27), where a low MATTR reflects
the repetitive, chant-like urgency of phrases such as "Calling all black
people." Conversely, Darwish’s lowest-MATTR poem, *I See My Ghost Coming
from Afar* (أرى شبحي قادماً من بعيد) (0.79), uses the deliberate
repetition of "I gaze upon" (أطل) as a rhetorical strategy common in
Arabic poetic traditions.

Overall, the high scores in poems like *Ka’Ba* and *I Am From There*
(أنا من هناك) suggest that resistance is not constructed through a fixed
or repetitive vocabulary. Instead, both poets employ a wide range of
images, cultural references, and identity markers to articulate
resistance. The fact that variation within each poet’s work is greater
than the difference between the two suggests that lexical diversity is
driven by individual poem structure rather than linguistic or stylistic
constraints. As noted in Section 1.2, these scores must be interpreted
as comparative within the corpus, as Arabic’s morphological complexity
inherently inflates TTR relative to English.

## 3.3 Sparsity and DFM Analysis

While, Lexical Diversity measured the vocabulary variation within
individual poems, Document–Feature Matrix (DFM) analysis allows us to
examine vocabulary distribution across the corpus as a whole. With this
approach, we look how much vocabulary is shared between poems versus how
much is unique to individual poems through the measure of sparsity. This
metric is computed within each poet's sub-corpus, making it comparable
across languages.

```{r sparsity-compute}
# Overall sparsity per poet
darwish_sparsity <- round(sparsity(darwish_dfm) * 100, 1)
baraka_sparsity  <- round(sparsity(baraka_dfm) * 100, 1)

# DFM dimensions
dfm_stats <- tibble(
  Poet = c("Mahmoud Darwish", "Amiri Baraka"),
  Documents = c(ndoc(darwish_dfm), ndoc(baraka_dfm)),
  Features  = c(nfeat(darwish_dfm), nfeat(baraka_dfm)),
  `Features per Doc` = round(c(nfeat(darwish_dfm) / ndoc(darwish_dfm),
                                nfeat(baraka_dfm) / ndoc(baraka_dfm))),
  `Sparsity (%)` = c(darwish_sparsity, baraka_sparsity)
)

kable(dfm_stats,
      col.names = c("Poet", "Documents", "Features", "Features per Doc", "Sparsity (%)"),
      caption = "Document-Feature Matrix comparison")
```

```{r sparsity-trim, fig.height=5}
# Sparsity at different trim levels
trim_levels <- seq(0, 0.5, by = 0.05)

sparsity_by_trim <- map_dfr(trim_levels, function(min_df) {
  d_trim <- dfm_trim(darwish_dfm, min_docfreq = max(min_df, 0.01),
                     docfreq_type = "prop")
  b_trim <- dfm_trim(baraka_dfm, min_docfreq = max(min_df, 0.01),
                     docfreq_type = "prop")
  tibble(
    min_docfreq = min_df,
    Darwish_sparsity = ifelse(nfeat(d_trim) > 0, sparsity(d_trim), NA_real_),
    Baraka_sparsity  = ifelse(nfeat(b_trim) > 0, sparsity(b_trim), NA_real_),
    Darwish_features = nfeat(d_trim),
    Baraka_features  = nfeat(b_trim)
  )
})

sparsity_by_trim %>%
  select(min_docfreq, Darwish_sparsity, Baraka_sparsity) %>%
  pivot_longer(-min_docfreq, names_to = "poet", values_to = "sparsity") %>%
  mutate(poet = str_remove(poet, "_sparsity")) %>%
  ggplot(aes(x = min_docfreq, y = sparsity, color = poet)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(labels = percent_format()) +
  labs(title = "Sparsity vs. Minimum Document Frequency Threshold",
       subtitle = "How sparsity changes as rare terms are progressively removed",
       x = "Minimum Document Frequency (proportion of poems)",
       y = "Sparsity", color = "Poet") +
  theme_minimal(base_size = 13)
```

The plot shows how much of a poet's vocabulary is unique to a few poems
versus how much is shared across their entire work. On the far left, the
high starting points, above 90%, reveal that both Darwish's and Baraka'
poems have high sparsity. This confirms the finding from Lexical
Diversity analysis, that both authors use a rich and diverse language.
Darwish's DFM has a slightly higher sparsity (92.1%) than Baraka
(91.4%), meaning that a given word in Darwish's corpus is less likely to
appear across multiple poems. The stagnation of the graphs as the filter
moves to the right along the x-axis, shows that once unique,
poem-specific words are removed, only a very small core of shared
thematic words remains.

## 3.4 Word Frequency Analysis

Word frequency analysis allows us to identify what the most frequent
words across corpora are and can reveal the core thematic vocabulary.

Since Arabic and English words cannot be compared directly, we display
each poet's top words separately in a side-by-side visualization.

```{r word-freq}
# Darwish word frequencies (using normalized Arabic stopwords)
darwish_words <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% ar_stops,
         str_length(word) > 1) %>%
  count(word, sort = TRUE)

# Baraka word frequencies (using unified English stopwords)
baraka_words <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% en_stops,
         str_length(word) > 2) %>%
  count(word, sort = TRUE)
```

```{r freq-barplot, fig.height=7}
top_d <- darwish_words %>% slice_max(n, n = 15) %>%
  mutate(poet = "Mahmoud Darwish")
top_b <- baraka_words %>% slice_max(n, n = 15) %>%
  mutate(poet = "Amiri Baraka")

bind_rows(top_d, top_b) %>%
  ggplot(aes(x = reorder_within(word, n, poet), y = n, fill = poet)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poet, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Top 15 Most Frequent Words by Poet",
       subtitle = "After stopword removal",
       x = NULL, y = "Count") +
  theme_minimal(base_size = 13)
```

```{r wordclouds, fig.height=10, fig.width=16}
wc_data <- bind_rows(
  darwish_words %>% slice_max(n, n = 80) %>% mutate(poet = "Mahmoud Darwish"),
  baraka_words  %>% slice_max(n, n = 80) %>% mutate(poet = "Amiri Baraka")
)

ggplot(wc_data, aes(label = word, size = n, color = n)) +
  geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle") +
  scale_size_area(max_size = 24) +
  scale_color_gradient(low = "darkgreen", high = "darkred") +
  facet_wrap(~poet, scales = "free") +
  theme_minimal(base_size = 14) +
  labs(title = "Word Clouds: Darwish vs. Baraka",
       subtitle = "Top 80 words after stopword removal")
```

Darwish's most frequent words are "look/gaze out/upon" (اطل), "and I"
(وانا), "my father" (ابي), "the sea" (البحر), and "and you" (وانت). This
vocabulary is rooted in the self, landscape, and family. It reflects
poetry of personal witness and memory, where Darwish positions himself
as the speaker and orientates himself in relation to land, family, and
an adressed other. The frequency of "and I" and "and you" suggests that
Darwish's resistance is expressed through intimate relationship and
personal testimony rather than collective declaration.

In contrast, Baraka’s most frequent words are “black,” “people,”
“killed,” “poems,” and “world,” reflecting a vocabulary centered on
collective identity, political struggle, and oppression. One interesting
finding from Baraka's word cloud is the simultaneous presence of both
'black' and 'white' as frequent words. This suggests that Baraka's
vocabulary defines Blackness in opposition to whiteness, reflecting the
systemic racial contrast at the heart of his political project.
Resistance here is not expressed in isolation but always in relation to
the oppressive structure it pushes against. His language addresses
communities, systems of power, and violence rather than the individual
speaker.

Particularly interesting is the cluster of violence-related words,
"killed", "dead", "shot", and "fire", appearing alongside "survive".
This suggests that resistance in Baraka's work is fundamentally about
endurance and survival in the face of physical and systemic threat
rather than abstract political struggle.

Both poets root their resistance in geopgraphic terms. Baraka uses words
as "land" and "america" frequently, positioning the resistance and black
struggle within and against a specific national territory. Darwish's
most frequent geographic words, "الارض"(the land/earth), "البحر" (the
sea), "السماء" (the sky), are less nationally specific and instead work
as symbolic landscape elements. This reflects the memories of his
homeland as remembered through the nature and environment rather than
politics and oppression.

Overall, Darwish's poetry is organized around the speaker, memory, and
landscape, while Baraka's poetry is organized around collective
identity, political struggle, and historical violence. Both are
resistance vocabularies, but they operate through fundamentally
different relationships between the poet, the people, and the land.

## 3.5 Bigram Network Graph

While word frequency shows what words a poet uses most, bigrams reveal
how those words are combined, exposing the rhetorical structures that
word counts alone cannot capture. As with word frequency, bigrams are
analyzed separately per poet.

```{r bigrams}
# Darwish bigrams (normalized Arabic stopwords)
darwish_bigrams <- darwish_poems %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% ar_stops, !word2 %in% ar_stops,
         str_length(word1) > 1, str_length(word2) > 1) %>%
  count(word1, word2, sort = TRUE)

# Baraka bigrams (unified English stopwords)
baraka_bigrams <- baraka_poems %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% en_stops, !word2 %in% en_stops,
         str_length(word1) > 2, str_length(word2) > 2) %>%
  count(word1, word2, sort = TRUE)

cat("Darwish unique bigrams:", nrow(darwish_bigrams), "\n")
cat("Baraka unique bigrams:", nrow(baraka_bigrams), "\n")
```

```{r bigram-table}
kable(head(darwish_bigrams, 10),
      col.names = c("Word 1", "Word 2", "Count"),
      caption = "Top 10 Darwish bigrams (after stopword removal)")
kable(head(baraka_bigrams, 10),
      col.names = c("Word 1", "Word 2", "Count"),
      caption = "Top 10 Baraka bigrams (after stopword removal)")
```

```{r bigram-network, fig.height=8, fig.width=14}
# Network plotting function
plot_bigram_network <- function(bigram_counts, title_text,
                                 n_edges = 25, min_n = 2) {
  graph_data <- bigram_counts %>%
    filter(n >= min_n) %>%
    slice_max(n, n = n_edges)

  if (nrow(graph_data) == 0) {
    # If no bigrams meet threshold, lower it
    graph_data <- bigram_counts %>% slice_max(n, n = n_edges)
  }

  graph <- graph_from_data_frame(graph_data)

  ggraph(graph, layout = "fr") +
    geom_edge_link(
      aes(edge_alpha = n),
      edge_width = 0.4,
      show.legend = FALSE,
      arrow = arrow(type = "closed", length = unit(0.12, "inches")),
      end_cap = circle(5, "mm")
    ) +
    geom_node_point(color = "steelblue", size = 5, alpha = 0.8) +
    geom_node_text(aes(label = name), 
                   vjust = 1.2, 
                   hjust = 0.5,
                   size = 3.5,
                   check_overlap = TRUE,
                   repel = TRUE) +
    theme_void(base_size = 13) +
    labs(title = title_text)
}

# Side-by-side networks
set.seed(42)
p_darwish <- plot_bigram_network(darwish_bigrams,
                                  "Darwish: Bigram Network",
                                  n_edges = 30, min_n = 2)
p_baraka  <- plot_bigram_network(baraka_bigrams,
                                  "Baraka: Bigram Network",
                                  n_edges = 30, min_n = 2)

p_darwish + p_baraka +
  plot_annotation(
    title = "Bigram Network Graphs: Darwish vs. Baraka",
    subtitle = "Nodes = words, edges = consecutive word pairs (after stopword removal)",
    theme = theme(plot.title = element_text(size = 16, face = "bold"))
  )
```

Due to Arabic morphology, bigrams in Arabic often represent larger
semantic phrases than in English. Because conjunctions and pronouns are
frequently attached to words, Arabic bigrams can capture full poetic
expressions, whereas English bigrams more often reflect simple
collocations such as noun phrases. This difference should be taken into
account when comparing bigram patterns across languages.

Baraka's bigrams cluster around themes of collective identity, political
accusation, and historical reference. His top bigram is "Black people"
functions as both a subject and a collective call, supported by its
connection to "Calling" in the network graph. This is reinforced by
"survive blackman / blackman survive," which creates a chant-like
inversion mirroring the repetitive style of *SOS*. Survival is not only
expressed, but repeated and reversed, suggesting that endurance under
oppression becomes a rhetorical strategy in itself.

Baraka’s rhetorical strategy of direct accusation is further evidenced
by "central bank" and "evil word." In the controversial *Somebody Blew
Up America*, bigrams like "4000 Israeli" and "1944 Minnesota" move the
poetry from abstract struggle to specific, historical indictments. By
listing institutional targets, Baraka names the enemy directly. Even his
metaphorical bigrams, such as "owl exploding," serve to intensify this
literal, confrontational register.

Conversely, Darwish’s bigrams are less centered on direct accusation and
instead cluster around imagery, land, and collective identity. Three of
Darwish's top bigrams, "اريد اطل" , "اطل كشرفه" and "كشرفه بيت" , form a
single continuous phrase "I want to look out like a balcony of a house".
The line appears five times in أرى شبحي قادماً من بعيد” (*I See My Ghost
Coming from Afar*), which shows Darwish's use of anaphoric repetition as
a rhetorical strategy for emphasizing key phrases and create rhythm.
Along with bigrams such as “بعيدة السماء” (the sky is far), “وانا ارضها”
(her land and I), and “اينما تذهب” (wherever you go), this imagery
points to distance, movement, and attachment to land. The speaker is
positioned as an observer looking toward a home or landscape that is no
longer fully accessible due to displacement and occupation, reflecting
themes of exile and disrupted belonging.

The remaining bigrams clusters around a few broader themes. Phrases like
“فكر بغيرك” (think of others) and “جميعهم أهلي” (all of them are my
people) point to a collective sense of identity and solidarity, where
resistance is framed through shared experience rather than direct
confrontation. This suggests that Darwish sees resistance not only by
personal experience but in a broader sense of community shaped by
displacement and occupation.

While Baraka builds resistance through overt confrontation, Darwish
often frames it through solidarity and the assertion of identity.
Phrases like "think of others" (فكر بغيرك) and "all of them are my
people" (جميعهم أهلي) point to a collective identity shaped by shared
displacement. Even the famous "Record! I am angry" (سجل تغضب) from
*Identity Card* directs the question of anger back onto the oppressor.
Rather than naming an enemy, Darwish resists by asserting the "I am" in
the face of erasure. Structurally, Baraka builds resistance by naming
the external enemy, while Darwish builds it by indirectly constructing
the scene of loss.

## 3.6 TF-IDF: Distinctive Words

TF-IDF moves the comparison to the poem level, asking not just how the
poets differ overall, but whether those differences are consistent
across individual works or vary within each poet’s corpus.

```{r tfidf, fig.height=8, fig.width=12}
# TF-IDF per poem within each poet
darwish_tfidf <- darwish_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% ar_stops, str_length(word) > 1) %>%
  count(poem, word, sort = TRUE) %>%
  bind_tf_idf(word, poem, n) 

baraka_tfidf <- baraka_poems %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% en_stops, str_length(word) > 2) %>%
  count(poem, word, sort = TRUE) %>%
  bind_tf_idf(word, poem, n)

# Top TF-IDF words per poet (across all poems)
top_tfidf_d <- darwish_tfidf %>%
  arrange(desc(tf_idf)) %>%
  distinct(word, .keep_all = TRUE) %>%
  slice_max(tf_idf, n = 15) %>%
  mutate(poet = "Mahmoud Darwish")

top_tfidf_b <- baraka_tfidf %>%
  arrange(desc(tf_idf)) %>%
  distinct(word, .keep_all = TRUE) %>%
  slice_max(tf_idf, n = 15) %>%
  mutate(poet = "Amiri Baraka")

bind_rows(top_tfidf_d, top_tfidf_b) %>%
  ggplot(aes(x = reorder_within(word, tf_idf, poet),
             y = tf_idf, fill = poet)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poet, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Most Distinctive Words by Poet (TF-IDF)",
       subtitle = "Words that are uniquely important to individual poems",
       x = NULL, y = "TF-IDF Score") +
  theme_minimal(base_size = 13)
```

**TF-IDF per poem**

We selected eight poems representing key themes such as political
resistance, violence, identity, exile, memory, cultural identity, and
philosophical introspection. This ensures that the visualization
reflects thematic diversity across both poets rather than only the
highest-scoring TF-IDF texts.

```{r tfidf-per-poem, fig.height=9, fig.width=13}
# Top TF-IDF words per poem -- Baraka (4 poems)
baraka_tfidf %>%
  filter(poem %in% c("Somebody Blew Up America", "Ka'Ba",
                      "Preface to a Twenty Volume Suicide Note",
                      "An Agony. As Now.")) %>%
  group_by(poem) %>%
  slice_max(tf_idf, n = 8) %>%
  ungroup() %>%
  ggplot(aes(x = reorder_within(word, tf_idf, poem),
             y = tf_idf, fill = poem)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poem, scales = "free", ncol = 2) +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Baraka: Most Distinctive Words per Poem (TF-IDF)",
       subtitle = "Words that define each poem's unique vocabulary",
       x = NULL, y = "TF-IDF") +
  theme_minimal(base_size = 12)
```

```{r td-idf-darwish, fig.height=9, fig.width=13}
# Top TF-IDF words per poem -- Darwish (4 poems)
darwish_tfidf %>%
  filter(poem %in% c("بطاقة هوية", "ريتا والبندقية",
                      "إلى أمي",
                      "أرى شبحي قادماً من بعيد")) %>%
  group_by(poem) %>%
  slice_max(tf_idf, n = 8) %>%
  ungroup() %>%
  ggplot(aes(x = reorder_within(word, tf_idf, poem),
             y = tf_idf, fill = poem)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~poem, scales = "free", ncol = 2) +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Darwish: Most Distinctive Words per Poem (TF-IDF)",
       subtitle = "Words that define each poem's unique vocabulary",
       x = NULL, y = "TF-IDF") +
  theme_minimal(base_size = 12)
```

TF-IDF reveals that for both poets, resistance is constructed through a
divided linguistic strategy rather than one single theme. Baraka’s top
words, "calling," "survive," "urgent," and "shot", are directly
connected to the themes of collective struggle and systemic oppression
identified earlier.

Looking at Baraka’s selected poems, his high scores in Anger and Fear
are built through a unique vocabulary of accusation in *Somebody Blew Up
America*, where terms like "killed" and "devil" serve as rhythmic
anchors. In contrast, *Preface to a Twenty Volume Suicide Note* reveals
a personal and philosophical dimension of resistance. Here, the use of
"stars," "night," and "daughter’s" mirrors an inward-looking struggle
with existential despair rather than outward conflict. Similarly, *An
Agony. As Now* utilizes a graphic vocabulary of "pain" and "flesh" to
describe the physical toll of oppression, while *Ka’Ba* shifts toward
the affirmation of identity through references to "African" and the
collective "walk." This shows that Baraka has a "political vocabulary"
he shares across many poems, but a "personal vocabulary" he keeps
isolated in specific, intimate works.

Darwish’s top words, "اطل" (look out/gaze), "ريتا" (Rita), "سجل"
(record), and "بغيرك" (others), reinforce the themes highlighted in his
bigrams. "اطل" and "ريتا" evoke memory and personal witness, while
"بغيرك" is connected to solidarity and collective identity. Together,
these terms show how Darwish’s vocabulary combines introspection and
memory to express resistance in a symbolic manner.

In *I See My Ghost Coming from Afar*, the exceptionally high score for
"اطل" (look out/gaze) confirms the structural use of anaphora. In
*Identity Card*, the terms "سجل" (record) and "عربي" (Arab) anchor the
poem’s confrontation with the occupier, asserting an identity rooted in
"roots" and "anger." These terms formed the rhythmic basis of the poem
as a protest song. Furthermore, the resistance in *Rita and the Rifle*
is highlighted by the unique presence of "Rita" and "rifle," blending
intimacy with militaristic reality. Finally, in *To My Mother*, domestic
terms like "bread" (**خبز**) and "coffee" (**قهوة**) serve as cultural
artifacts of memory as a defense against the erasure of exile.

Across these works, TF-IDF reveals that Baraka divides sharply between a
confrontational political vocabulary and an intimate one, Darwish's
distinctive terms are more unified. Whether a bureaucratic command or a
domestic memory, they all perform the same function, insisting on the
presence of a Palestinian reality that exile and occupation work to
erase.

# 4.0 Limitations

1.  **Corpus size**: With 14 poems per poet, this is a small curated
    sample by text mining standards and can introduce sample bias.
    Results therefore are exploratory and should not be generalized to
    the full body of either poet's work. The small corpus also precluded
    unsupervised topic modelling with LDA and thematic clusters were
    therefore identified through interpretive analysis of word
    frequency, bigrams, and TF-IDF.

2.  **Arabic morphology**: As discussed in Section 1.2, Arabic's
    inflectional system produces multiple surface forms from a single
    root. Without lemmatization, each form is counted separately, which
    likely inflates Darwish's TTR and sparsity scores and reduces NRC
    match rates relative to Baraka. Morphological differences also
    affect bigram analysis, as Arabic bigrams often represent larger
    semantic units than English bigrams.

3.  **NRC lexicon coverage**: The multilingual NRC lexicon relies on
    machine translation for non-English entries. Some Arabic emotion
    assignments may be imprecise due to translation artifacts or
    cultural differences in emotional connotation.

# 5. Conclusion

This project analyzed 28 poems through six text-mining techniques,
moving from basic tokenization to sentiment and network analysis. Across
these metrics, a consistent pattern emerges. While Darwish and Baraka
share a dominant emotional register of sadness and fear and comparable
lexical diversity, they diverge fundamentally in their rhetorical
construction of resistance.

The data confirms that Baraka expresses resistance through
confrontation. His resistance is collective and confrontational, built
on a structural anchor of terms like "Black people" and "killed." His
bigram networks and TF-IDF scores reveal a language addressed to a
community and an enemy, where the unit of resistance is a systemic
accusation against institutional targets like the "central bank".
Structurally, Baraka builds resistance outward toward the political
sphere.

In contrast, Darwish expresses ressitance through symbolic means. His
resistance is intimate and symbolic, using a vocabulary of "home" with
domestic artifacts like "bread" (خبز) and "coffee" (قهوة) and natural
imagery like the "sea" (بحر). These terms function as a defense against
the erasure of exile, affirming a Palestinian reality that exists
regardless of political occupation. Structurally, Darwish builds
resistance inward, using the individual speaker as a site of national
witness.

While Baraka’s corpus is characterized by emotional density and direct
systemic critique, Darwish’s corpus operates through lyrical
concentration, where memory and anaphora serve as forms of witness. This
suggests that while the experience of oppression is universal, the
linguistic responses to it are deeply rooted in specific cultural and
political traditions. Future research using larger corpora and
lemmatized Arabic text could further explore how these distinct
"resistance vocabularies" evolve over time or across different
translations.

# References

-   Benoit, K. et al. (2018). quanteda: An R package for the
    quantitative analysis of textual data. *Journal of Open Source
    Software*, 3(30), 774.
-   Covington, M. A. & McFall, J. D. (2010). Cutting the Gordian Knot:
    The Moving-Average Type-Token Ratio (MATTR). *Journal of
    Quantitative Linguistics*, 17(2), 94-100.
-   Mohammad, S. M. & Turney, P. D. (2013). Crowdsourcing a Word-Emotion
    Association Lexicon. *Computational Intelligence*, 29(3), 436-465.
-   Silge, J. & Robinson, D. (2017). *Text Mining with R: A Tidy
    Approach*. O'Reilly.
-   Welbers, K. et al. (2017). Text Analysis in R. *Communication
    Methods and Measures*, 11(4), 245-265.
-   Poetry Foundation. (n.d.). Mahmoud Darwish.
    <https://www.poetryfoundation.org/poets/mahmoud-darwish>
-   Poetry Foundation. (n.d.). Amiri Baraka.
    <https://www.poetryfoundation.org/poets/amiri-baraka>
-   Poem sources: aldiwan.net, poetsgate.com, poetryfoundation.org,
    poemhunter.com, Genius, becauseofthemwecan.com

------------------------------------------------------------------------

*Analysis conducted with R `r getRversion()`, using tidytext, quanteda,
and the NRC Emotion Lexicon.*

```{r session-info}
sessionInfo()
```
